---
title: "Kleb neonatal sepsis transmission estimates"
author: Erkison Odih
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        code_folding: hide
        df_print: paged
        code_download: true
---

```{r setup, include=FALSE, message=F, warning=F}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

<br>  

#### Packages and functions

<details>
<summary>Click to expand</summary> 
```{r load_packages, echo=F, message=F, warning=F}
library(tidyverse)
library(magrittr)
library(glue)
library(ggplot2)
library(plotly)
library(ggpubr)
library(patchwork)
library(ggnetwork)
library(igraph)
library(future)
library(meta)
library(metafor)
library(grid)
library(gtable)
library(sf)
```

```{r pkg_versions, class.source = 'fold-show'}
packageVersion("tidyverse")
packageVersion("magrittr")
packageVersion("glue")
packageVersion("ggplot2")
packageVersion("plotly")
packageVersion("ggpubr")
packageVersion("patchwork")
packageVersion("ggnetwork")
packageVersion("igraph")
packageVersion("future")
packageVersion("meta")
packageVersion("metafor")
packageVersion("grid")
packageVersion("gtable")
packageVersion("sf")
```

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}

```

```{r functions, message=F, warning=F}
# load functions
source('https://raw.githubusercontent.com/klebgenomics/transmission_estimator/main/src/functions/cluster_functions.R')
source('https://raw.githubusercontent.com/klebgenomics/transmission_estimator/main/src/functions/sensitivity_functions.R')
source('https://raw.githubusercontent.com/klebgenomics/transmission_estimator/main/src/functions/comparison_functions.R')
source('https://raw.githubusercontent.com/klebgenomics/transmission_estimator/main/src/functions/plotting_functions.R')
source('https://raw.githubusercontent.com/klebgenomics/transmission_estimator/main/src/functions/load_data_functions.R')
source("misc_functions.R")


custom_plots_theme <- ggplot2::theme(
    axis.text = ggplot2::element_text(size = 12),
    axis.title = ggplot2::element_text(size = 14),
    legend.text = ggplot2::element_text(size = 12),
    legend.title = ggplot2::element_text(size = 14),
    plot.title = element_text(size = 14)
)
```
</details>

<br>

#### Set up

```{r data_and_vars, class.source = 'fold-show'}
seed <- 24

# Clustering options / thresholds
snp_distance_threshold <- 10 # SNPs
temporal_distance_threshold <- 4 # weeks
geo_column <- 'Site'

# Sensitivity options
sens_snp_range <- seq(0, 25, by = 1)
sens_temp_range <- c(1:52)

# Sensitivity plot options
sens_temp_dist_vals <- c(1,2,4,8,52)

# figures and tables dir
figs_dir <- file.path("./figures_test/")
fs::dir_create(figs_dir)
tables_dir <- file.path("./tables_test/")
fs::dir_create(tables_dir)

index_cols <- c("Transmission proportion"="navyblue", "Cluster proportion"="#8b0000")
# res_colors <- c(`0: ESBL-, Carb-`="lightblue", `1: ESBL+, Carb-`="orange", 
#                 `2: Carb+`="red", `3: Carb+, Col+`="black")
res_colors <- c(`0: ESBL-, Carb-` ="#FEE0D2", 
                `1: ESBL+, Carb-` ="#FC9272", 
                `2: Carb+`        ="#EF3B2C", 
                `3: Carb+, Col+`  ="#99000D")
region_order <- c("Eastern Africa", "Western Africa", "Southern Africa", "Southern Asia")

```

<br>

## Studies  

#### BARNARDS
<details>
<summary>Click to expand</summary> 
```{r BARNARDS, fig.width=9, fig.height=9}
# load data
BARNARDS_snp_and_epi_data <- read_csv("data/BARNARDS/BARNARDS_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
BARNARDS_data <- read_csv("data/BARNARDS/BARNARDS_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
BARNARDS_epi_snp_clusters <- 
    get_clusters(BARNARDS_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=BARNARDS_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
BARNARDS_sens_df <- get_cluster_sensitivity(BARNARDS_snp_and_epi_data, BARNARDS_data,
                                            sens_snp_range, sens_temp_range)
```
</details>


<br>


#### GBS
<details>
<summary>Click to expand</summary> 
```{r GBS, fig.width=9, fig.height=9}
# load data
GBS_snp_and_epi_data <- read_csv("data/GBS/GBS_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
GBS_data <- read_csv("data/GBS/GBS_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
GBS_epi_snp_clusters <- 
    get_clusters(GBS_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=GBS_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
GBS_sens_df <- get_cluster_sensitivity(GBS_snp_and_epi_data, GBS_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>


#### KWTRP
<details>
<summary>Click to expand</summary> 
```{r KWTRP, fig.width=9, fig.height=9}
# load data
KWTRP_snp_and_epi_data <- read_csv("data/KWTRP/KWTRP_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
KWTRP_data <- read_csv("data/KWTRP/KWTRP_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
KWTRP_epi_snp_clusters <- 
    get_clusters(KWTRP_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=KWTRP_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
KWTRP_sens_df <- get_cluster_sensitivity(KWTRP_snp_and_epi_data, KWTRP_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>


#### NeoBAC
<details>
<summary>Click to expand</summary> 
```{r NeoBAC, fig.width=9, fig.height=9}
# load data
NeoBAC_snp_and_epi_data <- read_csv("data/NeoBAC/NeoBAC_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
NeoBAC_data <- read_csv("data/NeoBAC/NeoBAC_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
NeoBAC_epi_snp_clusters <- 
    get_clusters(NeoBAC_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=NeoBAC_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
NeoBAC_sens_df <- get_cluster_sensitivity(NeoBAC_snp_and_epi_data, NeoBAC_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>

#### SPINZ
<details>
<summary>Click to expand</summary> 
```{r SPINZ, fig.width=9, fig.height=9}
# load data
SPINZ_snp_and_epi_data <- read_csv("data/SPINZ/SPINZ_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
SPINZ_data <- read_csv("data/SPINZ/SPINZ_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
SPINZ_epi_snp_clusters <- 
    get_clusters(SPINZ_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=SPINZ_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
SPINZ_sens_df <- get_cluster_sensitivity(SPINZ_snp_and_epi_data, SPINZ_data,
                                            sens_snp_range, sens_temp_range)
```
</details>
<br>


#### BabyGERMS
<details>
<summary>Click to expand</summary> 
```{r BabyGERMS, fig.width=9, fig.height=9}
# load data
BabyGERMS_snp_and_epi_data <- read_csv("data/BabyGERMS/BabyGERMS_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
BabyGERMS_data <- read_csv("data/BabyGERMS/BabyGERMS_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
BabyGERMS_epi_snp_clusters <- 
    get_clusters(BabyGERMS_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=BabyGERMS_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
BabyGERMS_sens_df <- get_cluster_sensitivity(BabyGERMS_snp_and_epi_data, BabyGERMS_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>


#### MLW
<details>
<summary>Click to expand</summary> 
```{r MLW, fig.width=9, fig.height=9}
# load data
MLW_snp_and_epi_data <- read_csv("data/MLW/MLW_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
MLW_data <- read_csv("data/MLW/MLW_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
MLW_epi_snp_clusters <- 
    get_clusters(MLW_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=MLW_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
MLW_sens_df <- get_cluster_sensitivity(MLW_snp_and_epi_data, MLW_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>


#### NIMBI
<details>
<summary>Click to expand</summary> 
```{r NIMBI, fig.width=9, fig.height=9}
# load data
NIMBI_snp_and_epi_data <- read_csv("data/NIMBI/NIMBI_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
NIMBI_data <- read_csv("data/NIMBI/NIMBI_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
NIMBI_epi_snp_clusters <- 
    get_clusters(NIMBI_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=NIMBI_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
NIMBI_sens_df <- get_cluster_sensitivity(NIMBI_snp_and_epi_data, NIMBI_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>


#### DH
<details>
<summary>Click to expand</summary> 
```{r DH, fig.width=9, fig.height=9}
# load data
DH_snp_and_epi_data <- read_csv("data/DH/DH_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
DH_data <- read_csv("data/DH/DH_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
DH_epi_snp_clusters <- 
    get_clusters(DH_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=DH_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
DH_sens_df <- get_cluster_sensitivity(DH_snp_and_epi_data, DH_data,
                                            sens_snp_range, sens_temp_range)
```
</details>

<br>

#### MBIRA
<details>
<summary>Click to expand</summary> 
```{r MBIRA, fig.width=9, fig.height=9}
# load data
MBIRA_snp_and_epi_data <- read_csv("data/MBIRA/MBIRA_pairwise_snpepi_data.csv", show_col_types=F) %>% 
    mutate(across(c(iso1, iso2), \(x) as.character(x)))
MBIRA_data <- read_csv("data/MBIRA/MBIRA_data.csv", show_col_types=F) %>% mutate(id = as.character(id))

# get cluster graph based on snp dist, temporal dist, and shared location
MBIRA_epi_snp_clusters <- 
    get_clusters(MBIRA_snp_and_epi_data, genet_dist=snp_distance_threshold,
                 temp_dist=temporal_distance_threshold, metadata=MBIRA_data) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster)))

# Sensitivity
MBIRA_sens_df <- get_cluster_sensitivity(MBIRA_snp_and_epi_data, MBIRA_data,
                                            sens_snp_range, sens_temp_range)
```
</details>


<br><br>

## All data
```{r merge_studies, message=FALSE, warning=FALSE}
# merge clusters data
epi_snp_clusters <- bind_rows(
    BabyGERMS_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)),
    BARNARDS_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)), 
    DH_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)),
    GBS_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)), 
    KWTRP_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)),
    MBIRA_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)),
    MLW_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)),
    NeoBAC_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)), 
    NIMBI_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character)),
    SPINZ_epi_snp_clusters %>% mutate(across(everything(), .fns = as.character))
)

epi_snp_clusters <- readr::type_convert(epi_snp_clusters) %>% 
    mutate(Region = countrycode::countrycode(Country, "country.name", "region23")) %>%
    mutate(Region = factor(Region, levels=region_order)) %>% 
    mutate(Study_lab = case_when(
        Study == "BabyGERMS" ~ "Baby GERMS-SA",
        Study == "GBS" ~ "GBS-COP",
        Study == "NIMBI" ~ "NIMBI-plus",
        TRUE ~ Study
    )) %>% 
    mutate(coded_site = Site) %>% # ori Sites already coded so redundant; leave in or change use of coded_site throughout
    mutate(Country_Study = paste0(Country, ": ", Study_lab)) %>% 
    mutate(Country_Study_Site=paste0(Country_Study, ": ", coded_site)) %>%
    mutate(Study_Site=paste0(Study_lab, ": ", coded_site)) %>% 
    mutate(O_group = case_when(
        grepl("^O1a", O_type) ~ "O1",
        grepl("^O2a", O_type) ~ "O2",
        grepl("^O3", O_type) ~ "O3",
        TRUE ~ O_type
    ))

demog_table_1 <- epi_snp_clusters %>% 
    group_by(Region, Country, Study_lab, Site) %>% 
    reframe(n = n(),
            Period = ifelse(min(Year) == max(Year), 
                       as.character(min(Year)), 
                       paste0(min(Year), " - ", max(Year)))) %>% 
    arrange(Region, Country, Study_lab, desc(n)) %>% 
    rename("Num genomes" = "n") %>% 
    mutate("Num years" = if_else(grepl(" - ", Period), ">=2", "1"))
demog_table_1 %>% DT::datatable(., caption="Genome distribution")

# Studies
studies <- sort(unique(epi_snp_clusters$Study))
# Sites, ordered by num genomes
sites <- demog_table_1 %>% arrange(desc(`Num genomes`)) %>% pull(Site)

# overall raw proportions
overall_cluster_prop <- get_cluster_estimates(epi_snp_clusters)$cluster_prop
overall_cluster_prop
overall_transmission_prop <- get_cluster_estimates(epi_snp_clusters)$transmission_prop
overall_transmission_prop

```


### Summary stats
```{r summary}
epi_snp_clusters %>% count(Study) %>% arrange(desc(n)) %>% 
    DT::datatable(., caption="Studies")

epi_snp_clusters %>% count(Country) %>% arrange(desc(n)) %>% 
    DT::datatable(., caption="Countries")

epi_snp_clusters %>% count(Region) %>% arrange(desc(n)) %>% 
    DT::datatable(., caption="Regions")

all_clusters_info <-  get_cluster_info_nodates(epi_snp_clusters)

summarise_cluster_nodates(epi_snp_clusters) %>% 
    DT::datatable(., caption="Clusters summary")

# Exclude clusters of size 2 (possible transient transmission) and get estimates
epi_snp_clusters_of_gt2 <- epi_snp_clusters %>% 
    add_count(Cluster, name = "Cluster_size") %>% 
    mutate(Cluster = if_else(Cluster_size < 3, NA, Cluster)) %>% 
    select(-Cluster_size)

summarise_cluster_nodates(epi_snp_clusters_of_gt2) %>% 
    DT::datatable(., caption="Clusters summary (only clusters with ≥3 cases)")
```

<br> 

#### Regional distribution

```{r regional_distribution, warning=F, message=F, fig.width=12, fig.height=7}
region_colors <- setNames(c('#FFD92F', '#A95AA1', '#85C0F9', '#0F2080'),
                          unique(epi_snp_clusters$Region))

country_sample_count <- epi_snp_clusters %>% 
    group_by(Country) %>% 
    reframe(n_samples = n(), n_sites_per_country = n_distinct(Study_Site)) %>% 
    rename("country" = "Country")

get_breaks_and_mean <- function(vec, break_interval) {
  min_val <- min(vec)
  max_val <- max(vec)
  vals_mean <- mean(c(min_val, max_val))
  breaks <- seq(ceiling(min_val / break_interval) * break_interval, 
                floor(max_val / break_interval) * break_interval, 
                by = break_interval)
  return(list("vals_mean"=vals_mean, "breaks"=breaks))
}
ctry_smpl_brks_mean <- get_breaks_and_mean(country_sample_count$n_samples, 100)


# Sample distribution (Only Africa and Asia) --------------------------------
africa_and_asia <- rnaturalearth::ne_countries(scale="large", type="countries",
                                               continent=c("africa", "asia")) %>% 
    mutate(country = countrycode::countryname(name))
# sanity check non-matching countries
setdiff(country_sample_count$country, africa_and_asia$country)

# Sample distribution with sites shown (Only Africa and Asia) --------------------------------
site_coords <- read_csv("./data/site_coordinates.csv", show_col_types = F) %>% 
    left_join(demog_table_1 %>% select(Site, `Num genomes`, Period), by = "Site") 

site_sample_count_sf <- epi_snp_clusters %>% 
    group_by(Country, Site, coded_site, Study) %>% reframe(n_samples_per_site = n()) %>% 
    group_by(Country) %>% add_count(name="n_sites_per_country") %>% 
    rename("country" = "Country") %>%
    group_by(Study) %>% add_count(name="n_sites_per_study") %>%
    ungroup() %>%
    mutate(study_lab = if_else(n_sites_per_study==1, "Other (single site)", Study)) %>%
    left_join(africa_and_asia, by = c("country")) %>% 
    left_join(site_coords %>% select(Site, Long_city, Lat_city), by = "Site") %>% 
    mutate_at(vars(Long_city, Lat_city), as.numeric) %>% 
    # Jitter overlapping sites
    mutate(Lat_city = case_when(
        Study == "BARNARDS" & Site == "BABC" ~ Lat_city + 0.5,
        Study == "BARNARDS" & Site == "BAZAT" ~ Lat_city + 0.5,
        Study == "NeoBAC" & Site == "Mbagathi" ~ Lat_city - 0.25,
        TRUE ~ Lat_city
    )) %>% 
    arrange(desc(n_samples_per_site)) %>% 
    st_as_sf()


region_dist_wt_sites <- africa_and_asia %>% 
    left_join(country_sample_count, by = c("country")) %>% 
    ggplot() +
    geom_sf(aes(fill = n_samples)) +
    scale_fill_gradient2(low="#ffe599", mid="green", high="blue",
                         #low="yellow", mid="#ec7900", high="red",
                         midpoint=ctry_smpl_brks_mean[[1]],
                         na.value="white", breaks=ctry_smpl_brks_mean[[2]],
                         name="N isolates \nper country",
                         guide = guide_colorbar(order=1)) +
    geom_point(data=site_sample_count_sf,
               aes(x=Long_city, y=Lat_city, size=n_samples_per_site),
               fill="black", color="grey90", shape=21, stroke=0.15) +
    scale_size_continuous(range=c(1.5, 6), name = "N isolates \nper site",
                          guide = guide_legend(order=2, override.aes=list(stroke=0))) +
    ggthemes::theme_few() + custom_plots_theme +
    theme(axis.ticks = element_blank(), axis.text = element_blank(),
          axis.title = element_blank(), panel.border = element_blank())
```

<br> 

#### Year of isolation
```{r year_distribution, fig.width=7, fig.height=9}
year_plot2 <- epi_snp_clusters %>% 
    mutate(coded_site_num = as.numeric(gsub("Site |A|B", "", coded_site))) %>% 
    mutate(Region = factor(Region, levels=region_order)) %>% 
    arrange(factor(Region, levels=rev(region_order)), desc(coded_site_num)) %>%
    mutate(coded_site = factor(coded_site, levels = unique(coded_site))) %>% 
    add_count(coded_site, name="N_isolates") %>% 
    ggplot(aes(x=Year, y=coded_site)) +
    geom_jitter(aes(colour=Region), height=0.3, width=0.3) +
    geom_label(aes(y=coded_site, label=N_isolates, x=min(epi_snp_clusters$Year)-2), 
               fill="white", size=2.8, col="black") +
    scale_color_manual(values = region_colors, guide=guide_legend(override.aes=list(size=5))) +
    scale_x_continuous(breaks = scales::breaks_pretty()) +
    geom_hline(yintercept = seq(0.5, length(unique(epi_snp_clusters$coded_site))+0.5, 1), 
               linewidth=0.25, color="gray") +
    geom_vline(xintercept =seq(min(epi_snp_clusters$Year)-1.5, max(epi_snp_clusters$Year), 1), 
               linewidth=0.25, color="gray") +
    labs(x="Year", y=NULL) +
    theme_minimal() + custom_plots_theme +
    theme(axis.text.x = element_text(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          legend.position = "top", legend.title = element_blank(),
          legend.key.width = unit(5, "pt")) 

year_region_plot <- ggarrange(region_dist_wt_sites, year_plot2, ncol=1, 
                              heights = c(1,1.7), labels = "AUTO")
ggsave(file=file.path(figs_dir, "FigureS2_GeotemporalDistrib.png"), 
       plot=year_region_plot, width=9, height=12, bg="white")
ggsave(file=file.path(figs_dir, "FigureS2_GeotemporalDistrib.tif"), 
       plot=year_region_plot, width=9, height=12, bg="white")

```


<br><br>

### Sensitivity

```{r combined_sensitivity_plots, fig.width=15, fig.height=9}
all_sens_df <- lapply(studies, function(x) {
    study_lab <- epi_snp_clusters$Study_lab[epi_snp_clusters$Study == x][1]
    get(paste0(x, "_sens_df")) %>% mutate(Study = study_lab)
}) %>% bind_rows()

# legend
sens_clust_legend <- ggarrange(manual_sensitivity_legend2(mid_temp_val=sens_temp_dist_vals[3],
                           temp_range_1 = sens_temp_dist_vals[c(2,4)],
                           temp_range_2 = sens_temp_dist_vals[c(1,5)],
                           selected_dist_thresh=snp_distance_threshold,
                           selected_temp_thresh=temporal_distance_threshold)) 
sens_trans_legend <- ggarrange(manual_sensitivity_legend2(mid_temp_val=sens_temp_dist_vals[3],
                           temp_range_1 = sens_temp_dist_vals[c(2,4)],
                           temp_range_2 = sens_temp_dist_vals[c(1,5)],
                           selected_dist_thresh=snp_distance_threshold,
                           selected_temp_thresh=temporal_distance_threshold,
                           range_cols=c("white", "navyblue", "lightblue"))) 

# Cluster prop sensitivity
combined_sens_clust_plot <- (
    plot_sensitivity_SNP_vs_temp_range_2(
        all_sens_df, temp_dist_vals = sens_temp_dist_vals, 
        prop_var = "cluster_prop", y_title = "Proportion in clusters",
        plot_title = NULL, selected_dist_thresh=snp_distance_threshold,
        selected_temp_thresh=temporal_distance_threshold,
        facet_col = "Study", facet_rows = 3) |
    sens_clust_legend ) +
    plot_layout(widths=c(6.5,1), nrow=1)
combined_sens_clust_plot


# Transmission prop sensitivity
combined_sens_trans_plot <- (
    plot_sensitivity_SNP_vs_temp_range_2(
        all_sens_df, temp_dist_vals = sens_temp_dist_vals,
        prop_var = "transmission_prop", y_title = "Proportion due to transmission",
        plot_title = NULL, selected_dist_thresh=snp_distance_threshold,
        selected_temp_thresh=temporal_distance_threshold,
        range_cols=c("white", "navyblue", "lightblue"),
        facet_col = "Study", facet_rows = 3) |
    sens_trans_legend ) +
    plot_layout(widths=c(6.5,1), nrow=1)
combined_sens_trans_plot
```

<br><br>

```{r sensitivity_figure_manuscript, echo=F}
# # manuscript fig
# sens_fig_manuscript <- ggarrange(
#     combined_sens_clust_plot +
#         plot_annotation(title = "Cluster proportion estimates",
#                         theme=theme(plot.title = element_text(size=14, face="bold", hjust=0.1))),
#     combined_sens_trans_plot +
#         plot_annotation(title = "Transmission proportion estimates",
#                         theme=theme(plot.title = element_text(size=14, face="bold", hjust=0.1))),
#     ncol=1, labels=c("(A)", "(B)"), font.label=list(size=16, face="bold", color="black"))
# 
# ggsave(file=file.path(figs_dir, "FigureS4_ThresholdSensitivity.png"),
#        plot=sens_fig_manuscript, width=9, height=10)
# ggsave(file=file.path(figs_dir, "FigureS4_ThresholdSensitivity.tif"),
#        plot=sens_fig_manuscript, width=9, height=10)


ggsave(file=file.path(figs_dir, "FigureS4_ThresholdSensitivityClust.png"),
       plot=combined_sens_clust_plot, width=9, height=5)
ggsave(file=file.path(figs_dir, "FigureS4_ThresholdSensitivityClust.tif"),
       plot=combined_sens_clust_plot, width=9, height=5)

ggsave(file=file.path(figs_dir, "AppendixS1Fig1_ThresholdSensitivityClust.png"),
       plot=combined_sens_clust_plot, width=9, height=5)
ggsave(file=file.path(figs_dir, "AppendixS1Fig1_ThresholdSensitivityClust.tif"),
       plot=combined_sens_clust_plot, width=9, height=5)

ggsave(file=file.path(figs_dir, "AppendixS1Fig2_ThresholdSensitivityTrans.png"),
       plot=combined_sens_trans_plot, width=9, height=5)
ggsave(file=file.path(figs_dir, "AppendixS1Fig2_ThresholdSensitivityTrans.tif"),
       plot=combined_sens_trans_plot, width=9, height=5)
```

```{r sensitivity_diff_in_estimates, echo=F}

SNP_thresh_pair <- c(5, 25)
temp_thresh_pair <- c(2, 8)

# fnx to compute diff in estimates between two threshold_A points across all threshold_B points
get_sensitivity_estimates_diff <- function(sensitivity_df, threshold_A_points = c(5, 25),
         threshold_A="distance_threshold", threshold_B="temporal_threshold"){
    stopifnot(threshold_A_points %in% sensitivity_df[[threshold_A]])
    # useful col names for output
    if(threshold_A=="distance_threshold"){
        thresh_unit="SNPs"
    } else if (threshold_A=="temporal_threshold") {thresh_unit="weeks"}
    diff_clust_var <- paste("diff_clust_prop_at", threshold_A, 
                            paste(threshold_A_points, collapse="-"), thresh_unit, sep="_")
    diff_trans_var <- paste("diff_trans_prop_at", threshold_A,
                            paste(threshold_A_points, collapse="-"), thresh_unit, sep="_")
    # get difference between estimates at the two threshold A points 
    # for all threshold B points
    differences <- sensitivity_df %>% 
        filter(!!rlang::sym(threshold_A) %in% threshold_A_points) %>% 
        group_by(!!rlang::sym(threshold_B)) %>% 
        summarise(diff_clust = abs(max(cluster_prop) - min(cluster_prop)),
                  diff_trans = abs(max(transmission_prop) - min(transmission_prop))) %>% 
        rename(!!diff_clust_var := "diff_clust",
               !!diff_trans_var := "diff_trans") %>% 
        ungroup()
    return(differences)
}


# Differences in estimates within SNP range (5 - 25 SNPs) at 4 weeks -------
diff_within_SNP_range <- furrr::future_map_dfr(studies, ~{
    # get sensitivity data
    dataset_sens_df <- get(paste0(.x, "_sens_df")) %>% filter(temporal_threshold == 4)
    # calculate differences in estimates between two SNP thresholds
    get_sensitivity_estimates_diff(dataset_sens_df, threshold_A_points=SNP_thresh_pair,
                               threshold_A="distance_threshold", 
                               threshold_B="temporal_threshold") %>% 
        mutate(Study = .x)
})

# overall mean diff summary (in %)
diff_within_SNP_range %>% select(starts_with("diff_")) %>% 
    mutate(across(everything(), \(x) x*100)) %>% summary()


# Also calculate diff in clust prop within 0 - 5 SNPs ---------------------------
diff_SNP_0_5 <- furrr::future_map_dfr(studies, ~{
        dataset_sens_df <- get(paste0(.x, "_sens_df")) %>% filter(temporal_threshold == 4)
        get_sensitivity_estimates_diff(dataset_sens_df, c(0,5)) %>% mutate(Study = .x)
    }) %>% 
    select(starts_with("diff_clust_prop_at_distance_threshold")) %>% 
    mutate(across(everything(), \(x) x*100)) 
summary(diff_SNP_0_5)
# calculate diff in clust prop within 0 - 1 SNPs ---------------------------
diff_SNP_0_1 <- furrr::future_map_dfr(studies, ~{
        dataset_sens_df <- get(paste0(.x, "_sens_df")) %>% filter(temporal_threshold == 4)
        get_sensitivity_estimates_diff(dataset_sens_df, c(0,1)) %>% mutate(Study = .x)
    }) %>% 
    select(starts_with("diff_clust_prop_at_distance_threshold")) %>% 
    mutate(across(everything(), \(x) x*100)) 
summary(diff_SNP_0_1)
# calculate diff in clust prop within 4 - 5 SNPs ---------------------------
diff_SNP_4_5 <- furrr::future_map_dfr(studies, ~{
        dataset_sens_df <- get(paste0(.x, "_sens_df")) %>% filter(temporal_threshold == 4)
        get_sensitivity_estimates_diff(dataset_sens_df, c(4,5)) %>% 
            mutate(Study = .x)
    }) %>% 
    select(starts_with("diff_clust_prop_at_distance_threshold")) %>% 
    mutate(across(everything(), \(x) x*100)) 
summary(diff_SNP_4_5)



# Differences in estimates within temp dist range (2-8 weeks) at 10 SNPs ------
diff_within_temp_range <- furrr::future_map_dfr(studies, ~{
    # get sensitivity data
    dataset_sens_df <- get(paste0(.x, "_sens_df")) %>% filter(distance_threshold == 10)
    # calculate differences in estimates between two temp dist thresholds
    get_sensitivity_estimates_diff(dataset_sens_df, threshold_A_points=temp_thresh_pair,
                               threshold_A="temporal_threshold", 
                               threshold_B="distance_threshold") %>% 
        mutate(Study = .x)
})
mean_diff_within_temp_range_per_study <- diff_within_temp_range %>% 
    group_by(Study) %>% summarise_at(2:3, .funs = mean) %>% 
    ungroup()
# overall mean diff summary
diff_within_temp_range %>% select(starts_with("diff_")) %>% 
    mutate(across(everything(), \(x) x*100)) %>% summary()

# also calculate diff in clust prop within 0 - 1 weeks ---------------------------
diff_temp_1_2 <- furrr::future_map_dfr(studies, ~{
        dataset_sens_df <- get(paste0(.x, "_sens_df")) %>% filter(distance_threshold == 10)
        get_sensitivity_estimates_diff(
            dataset_sens_df, c(1,2), "temporal_threshold", "distance_threshold"
        ) %>% mutate(Study = .x)
    }) %>% 
    select(starts_with("diff_clust_prop_at_temp")) %>% 
    mutate(across(everything(), \(x) x*100)) 
summary(diff_temp_1_2)

```



<br><br>

```{r sensitivity_up_to_1000SNPs}
all_sens_df_1000SNPs <- lapply(studies, function(x) {
    study_lab <- epi_snp_clusters$Study_lab[epi_snp_clusters$Study == x][1]
    get_cluster_sensitivity(snp_and_epi_data = get(paste0(x, "_snp_and_epi_data")),
                            metadata = get(paste0(x, "_data")),
                            snp_range=c(seq(0,25, by=1), seq(30,1000,by=5)),
                            date_range=c(1,2,4,8,52)) %>%
        mutate(Study = study_lab)
}) %>% bind_rows()


# Cluster prop sensitivity 1000 SNPs
sens_clust_plot_1000SNPs <- (
    plot_sensitivity_SNP_vs_temp_range_2(
        all_sens_df_1000SNPs, temp_dist_vals = sens_temp_dist_vals,
        prop_var = "cluster_prop", y_title = "Proportion in clusters",
        plot_title = NULL, selected_dist_thresh=snp_distance_threshold,
        selected_temp_thresh=temporal_distance_threshold,
        facet_col = "Study", facet_rows = 3) +
        scale_x_continuous(trans = scales::pseudo_log_trans(base=10),
                           breaks = c(0,10,100,1000)) +
        labs(x = "Genetic distance threshold (log10 scale)") +
        theme(axis.text = element_text(size=12), axis.title = element_text(size=12),
              strip.text = element_text(size=12)) |
    sens_clust_legend ) +
    plot_layout(widths=c(6.5,1), nrow=1)

# Transmission prop sensitivity 1000 SNPs
sens_trans_plot_1000SNPs <- (
    plot_sensitivity_SNP_vs_temp_range_2(
        all_sens_df_1000SNPs, temp_dist_vals = sens_temp_dist_vals,
        prop_var = "transmission_prop", y_title = "Proportion due to transmission",
        plot_title = NULL, selected_dist_thresh=snp_distance_threshold,
        selected_temp_thresh=temporal_distance_threshold,
        range_cols=c("white", "navyblue", "lightblue"),
        facet_col = "Study", facet_rows = 3) +
        scale_x_continuous(trans = scales::pseudo_log_trans(base=10),
                           breaks = c(0,10,100,1000)) +
        labs(x = "Genetic distance threshold (log10 scale)") +
        theme(axis.text = element_text(size=12), axis.title = element_text(size=12),
              strip.text = element_text(size=12)) |
    sens_trans_legend ) +
    plot_layout(widths=c(6.5,1), nrow=1)

# Combine both figs
sens_fig_1000SNPs <- ggarrange(
    sens_clust_plot_1000SNPs +
        plot_annotation(title = "Cluster proportion estimates",
                        theme=theme(plot.title = element_text(size=14, face="bold", hjust=0.1))),
    sens_trans_plot_1000SNPs +
        plot_annotation(title = "Transmission proportion estimates",
                        theme=theme(plot.title = element_text(size=14, face="bold", hjust=0.1))),
    ncol=1, labels=c("(A)", "(B)"), font.label=list(size=16, face="bold", color="black"))

ggsave(file=file.path(figs_dir, "AppendixS1Fig3_ThresholdSensitivity1000SNPs.png"),
       plot=sens_fig_1000SNPs, width=10, height=10)
ggsave(file=file.path(figs_dir, "AppendixS1Fig3_ThresholdSensitivity1000SNPs.tif"),
       plot=sens_fig_1000SNPs, width=10, height=10)
```

<br><br>


### Pooled estimates meta-analysis (sites)

```{r meta_analysis_by_site, warning=F}
# sites
all_sites_df <- epi_snp_clusters %>% 
    group_by(Study_Site, coded_site, Region, Country, Study) %>% 
    reframe(n_isolates = n())

# combine cluster details all sites
all_sites_clusters_lst <- split(epi_snp_clusters, f = epi_snp_clusters$coded_site)
all_sites_clusters_summary <- all_sites_clusters_lst %>% 
    purrr::map_dfr(\(clusters_data) summarise_cluster_nodates(clusters_data) %>% 
                       pivot_wider(names_from=name, values_from=value),
                   .id = "Study")
colnames(all_sites_clusters_summary) <- c("site","n_isolates","clusters","n_isolates_clusters",
                                          "n_transmission_isolates", "median_cluster_size", "cluster_STs")
all_sites_clusters_summary %<>% 
    mutate(across(c(n_isolates, n_isolates_clusters, n_transmission_isolates), \(x) as.numeric(x))) %>%
    mutate(cluster_prop = n_isolates_clusters / n_isolates,
           transmission_prop = n_transmission_isolates / n_isolates) %>% 
    arrange(desc(cluster_prop)) %>% 
    left_join(all_sites_df %>% select(-c(n_isolates)), by = c("site" = "coded_site"))

# functions
get_metaprop_estimates <- function(meta_obj,
                                   prop_type=c("prop", "percent"), dec=2) {
    prop_type = match.arg(prop_type)
    trns = meta_obj$sm
    if (trns == "PFT") {n <- meta_obj$k.all.w} else {n <- NULL}
    if (prop_type == "prop") {mult <- 1} else {mult <- 100}
    # get estimates
    estimates <- data.frame(
        S = meta_obj$studlab,
        Proportion = round(meta::backtransf(meta_obj$TE, sm=trns, n=n)*mult, dec),  
        lower_CI = round(meta::backtransf(meta_obj$lower, sm=trns, n=n)*mult, dec), 
        upper_CI = round(meta::backtransf(meta_obj$upper, sm=trns, n=n)*mult, dec),
        pooled_Est = round(meta::backtransf(meta_obj$TE.random, sm=trns, n=n)*mult, dec),
        pooled_lower = round(meta::backtransf(meta_obj$lower.random, sm=trns, n=n)*mult, dec),
        pooled_upper = round(meta::backtransf(meta_obj$upper.random, sm=trns, n=n)*mult, dec)
    ) %>% as_tibble() %>% arrange(desc(Proportion))
    return(estimates)
}

plot_meta <- function(meta_obj, meta_df, events_lab, x_lab="Proportion") {
    meta_df %<>% 
        left_join(meta_obj$data %>% 
                      select(site, Region, Country, `Total isolates`=n_isolates,Events = .event), 
                  by = c("S" = "site")) %>% 
        mutate(Site_ctry = paste0(S, " (", Country, ")")) %>% 
        arrange(desc(Proportion)) %>% 
        mutate(Site_ctry = factor(Site_ctry, levels = unique(Site_ctry)))
    # Plot
    heterogeneity_lab <- paste0(
        "Heterogeneity: *I*<sup>2</sup> = ", sprintf('%.1f', meta_obj$I2*100),
        "% [", sprintf('%.1f', meta_obj$lower.I2*100), "%, ", 
        sprintf('%.1f', meta_obj$upper.I2*100), "%], tau<sup>2</sup> = ",
        sprintf('%.4f', meta_obj$tau2),
        ifelse(meta_obj$pval.Q[1] < 0.0001, ", *p* < 0.0001", glue(", *p* = {meta_obj$pval.Q[1]}"))
    )
    p.left <- meta_df %>% 
        ggplot(aes(y = Site_ctry)) +
        # Sites
        geom_text(aes(x=0, label = Site_ctry), hjust=0, size=2.5) +
        annotate("text", x=0, y=nrow(meta_df)+1, label="Site", hjust=0, fontface="bold", size=2.5) + 
        annotate("text", x=0, y=0, label="Random effects model", hjust=0, fontface="bold", size=2.5) +
        # N isolates
        geom_text(aes(x=1.7, label=`Total isolates`), hjust=0, size=2.5) +
        annotate("text", x=1.7, y=nrow(meta_df)+1, label="Total", hjust=0, fontface="bold", size=2.5) +
        annotate("text", x=1.7, y=0, label=sum(meta_obj$n), hjust=0, fontface="bold", size=2.5) +
        # N events
        geom_text(aes(x=2.2, label=Events), hjust=0, size=2.5) +
        annotate("text", x=2.2, y=nrow(meta_df)+1, label=events_lab, hjust=0, 
                 vjust=0.2, fontface="bold", size=2.5, lineheight = 0.75) +
        annotate("text", x=2.2, y=0, label=sum(meta_obj$event), hjust=0, fontface="bold", size=2.5) +
        # Prop
        geom_text(aes(x=3.1, label=sprintf("%.2f [%.2f, %.2f]", Proportion, lower_CI, upper_CI)), 
                  hjust=0, size=2.5) +
        annotate("text", x=3.1, y=nrow(meta_df)+1, label="Proportion\n[95% CI]", 
                 lineheight=0.75, vjust=0.2, hjust=0, fontface="bold", size=2.5) +
        annotate("text", x=3.1, y=0, 
                 label=sprintf("%.2f [%.2f, %.2f]", meta_df$pooled_Est[1], 
                               meta_df$pooled_lower[1], meta_df$pooled_upper[1]), 
                 hjust=0, fontface="bold", size=2.5) +
        # Heterogeneity
        ggtext::geom_richtext(aes(x=0, y = -1.4, label = heterogeneity_lab, family="sans"), 
                              data=meta_df[1,],
                              hjust=0, vjust = 0, size = 2.2, fill = NA, label.color = NA,
                              label.padding = grid::unit(rep(0, 4), "pt")) +
        coord_cartesian(ylim=c(-1,nrow(meta_df)+2), xlim=c(0, 4)) +
        theme_void() +
        theme(text = element_text(family = "Arial"))
    p.right <- meta_df %>%
        ggplot(aes(y = Site_ctry, x = Proportion, xmin = lower_CI, xmax = upper_CI)) +
        geom_vline(xintercept = meta_df$pooled_Est[1], linetype = "dashed") +
        # Individual site estimates (squares) and interv
        geom_pointrange(aes(fill = Region), shape = 22, size = 0.6, stroke=0.5) +
        scale_fill_manual(values = region_colors) +
        # Pooled estimate (diamond) and interval
        geom_pointrange(aes(y=0, x=pooled_Est, xmin=pooled_lower, xmax=pooled_upper),
                        data=meta_df[1,], shape = 22, size = 0.8, stroke = 0.8) +
        # clean
        scale_x_continuous(name = x_lab, breaks = seq(0, 1, 0.2)) +
        coord_cartesian(ylim=c(-1,nrow(meta_df)+2), xlim=c(0, 1)) +
        theme_classic() +
        theme(axis.line.y = element_blank(), axis.ticks.y= element_blank(),
              axis.text.y= element_blank(), axis.title.y= element_blank(),
              axis.text = element_text(size=7), axis.title = element_text(size=8),
              legend.text = element_text(size=7, margin = margin(l=0.5, r=0.5, unit = "mm")), 
              legend.title = element_text(size=7), legend.key.spacing.x = unit(10, "mm"))
    # p <- ggarrange(p.left, p.right, ncol = 2, widths=c(1.8,2), 
    #                align="h", legend="bottom", common.legend=T)
    
    return(list(p.left = p.left, p.right = p.right))
}
```

#### Cluster proportion meta-analysis

```{r cluster_prop_meta, fig.width=10, fig.height=8}
# Pooled estimates ---------------------------------------------------------------
# cluster proportion
set.seed(seed)
cluster_prop_meta <- meta::metaprop(
    n_isolates_clusters, n_isolates, studlab=site, data=all_sites_clusters_summary, 
    method="GLMM", method.ci="CP", # method.tau="DL",
    sm="PLOGIT")

summary(cluster_prop_meta)
cluster_prop_meta_df <- get_metaprop_estimates(cluster_prop_meta) 
cluster_prop_meta_plots <- plot_meta(cluster_prop_meta, cluster_prop_meta_df, 
                                      "Cluster\nisolates", "Cluster proportion")
cluster_prop_forest_plot <- ggarrange(plotlist = cluster_prop_meta_plots, 
                                      ncol = 2, widths=c(1.8,2),
                                      align="h", legend="bottom", common.legend=T)
cluster_prop_forest_plot
# ggsave(file=file.path(figs_dir, "meta_cluster_prop.png"),
#        plot=cluster_prop_forest_plot, width=14, height=8, bg='white')

```

<br>

#### Transmission proportion meta-analysis
```{r transm_prop_meta, fig.width=10, fig.height=8}
# Proportion due to transmission
set.seed(seed)
transmi_prop_meta <- meta::metaprop(
    n_transmission_isolates, n_isolates, studlab=site, data=all_sites_clusters_summary, 
    method="GLMM", method.ci="CP", # method.tau="DL",
    sm="PLOGIT")
summary(transmi_prop_meta)
transmi_prop_meta_df <- get_metaprop_estimates(transmi_prop_meta)
transmi_prop_meta_plots <- plot_meta(transmi_prop_meta, transmi_prop_meta_df, 
                                     "Transmitted\nisolates", "Transmission proportion")
transmi_prop_forest_plot <- ggarrange(plotlist = transmi_prop_meta_plots, 
                                      ncol = 2, widths=c(1.8,2),
                                      align="h", legend="bottom", common.legend=T)
transmi_prop_forest_plot
# ggsave(file=file.path(figs_dir, "meta_transmission_prop.png"), 
#        plot=transmi_prop_forest_plot, width=14, height=8, bg='white')


transmi_meta_report <- 
    get_metaprop_estimates(transmi_prop_meta, prop_type = "percent", dec=1) %>% 
    left_join(all_sites_clusters_summary %>% select(site, n_transmission_isolates, n_isolates),
              by = c("S" = "site")) %>% 
    mutate(report=paste(
        glue("{S} (n = {n_transmission_isolates}/{n_isolates};"),
        glue("{round(Proportion,0)}% [{round(lower_CI,0)}-{round(upper_CI,0)}%])")
    )) 
print(paste(
    glue("In {transmi_meta_report %>% filter(Proportion >= 50) %>% nrow()}"),
    glue("of the {length(unique(epi_snp_clusters$Site))} sites, at least half"),
    glue("of all cases were attributable to transmission. These included"),
    paste0(
        sub("(.*),", "\\1, and", 
            paste0(transmi_meta_report %>% filter(Proportion>=50) %>% pull(report), 
                   collapse = ", ")),
        "."),
    "Overall, at least a third of cases were attributable to transmission in",
    transmi_meta_report %>% filter(Proportion >= 33.3) %>% nrow(), "sites."
    ))
```


```{r meta_figures_manuscript}
combined_forest_plots_sites_v <- cluster_prop_forest_plot +
  ggtitle("(A) Cluster proportion") + 
  theme(plot.title.position = "plot", 
        plot.title = element_text(hjust = 0, size = 12, face = "bold",
                                  margin = margin(b = 5))) + 
  transmi_prop_forest_plot +
  ggtitle("(B) Transmission proportion") +
  theme(plot.title.position = "plot",
        plot.title = element_text(hjust = 0, size = 12, face = "bold",
                                  margin = margin(b=5, t=5))) +
  plot_layout(ncol = 1, heights = c(1,1))
ggsave(file=file.path(figs_dir, "Figure1_PooledEstimates.png"), 
       plot=combined_forest_plots_sites_v, width=2000, height=2600, units="px", bg='white')
ggsave(file=file.path(figs_dir, "Figure1_PooledEstimates.tif"), 
       plot=combined_forest_plots_sites_v, width=2000, height=2600, units="px", bg='white')
```

<br><br>

```{r multi_site_meta}
# Test heterogeneity in multisite studies using same protocols across sites

# BARNARDS
BARNARDS_cluster_prop_meta <- meta::metaprop(
    n_isolates_clusters, n_isolates, studlab=site, 
    data=all_sites_clusters_summary %>% filter(Study == "BARNARDS"), 
    method="GLMM", method.ci="CP", # method.tau="DL",
    sm="PLOGIT")
BARNARDS_cluster_prop_meta_df <- get_metaprop_estimates(BARNARDS_cluster_prop_meta)

BARNARDS_transmi_prop_meta <- meta::metaprop(
    n_transmission_isolates, n_isolates, studlab=site, 
    data=all_sites_clusters_summary %>% filter(Study == "BARNARDS"), 
    method="GLMM", method.ci="CP", # method.tau="DL",
    sm="PLOGIT")
BARNARDS_transmi_prop_meta_df <- get_metaprop_estimates(BARNARDS_transmi_prop_meta)

# MBIRA
MBIRA_cluster_prop_meta <- meta::metaprop(
    n_isolates_clusters, n_isolates, studlab=site, 
    data=all_sites_clusters_summary %>% filter(Study == "MBIRA"), 
    method="GLMM", method.ci="CP", # method.tau="DL",
    sm="PLOGIT")
MBIRA_cluster_prop_meta_df <- get_metaprop_estimates(MBIRA_cluster_prop_meta)

MBIRA_transmi_prop_meta <- meta::metaprop(
    n_transmission_isolates, n_isolates, studlab=site, 
    data=all_sites_clusters_summary %>% filter(Study == "MBIRA"), 
    method="GLMM", method.ci="CP", # method.tau="DL",
    sm="PLOGIT")
MBIRA_transmi_prop_meta_df <- get_metaprop_estimates(MBIRA_transmi_prop_meta)
```

<br>

### Pooled estimates meta-analysis (subgroup analysis)

#### Pooled estimates by region
```{r region_meta, fig.height=5, fig.width=10}
get_metaprop_subgroup_estimates <- function(meta_obj, trns=c("PLOGIT", "PFT"), 
                                            prop_type=c("prop", "percent"), dec=2) {
    trns = match.arg(trns)
    prop_type = match.arg(prop_type)
    if (trns == "PFT") {n <- meta_obj$k.all.w} else {n <- NULL}
    if (prop_type == "prop") {mult <- 1} else {mult <- 100}
    # get estimates
    subgroup_estimates <- data.frame(
        Subgroups = meta_obj$subgroup.levels,                     
        Proportion = round(meta::backtransf(meta_obj$TE.random.w, sm=trns, n=n)*mult, dec),  
        lower_CI = round(meta::backtransf(meta_obj$lower.random.w, sm=trns, n=n)*mult, dec), 
        upper_CI = round(meta::backtransf(meta_obj$upper.random.w, sm=trns, n=n)*mult, dec), 
        tau2 = meta_obj$tau2.w,                      
        I2 = meta_obj$I2.w,
        Q = meta_obj$Q.w
    ) %>% as_tibble()
    return(subgroup_estimates)
}

# factor - order regions by raw clust props
all_sites_clusters_summary %<>% arrange(desc(cluster_prop)) %>% 
    mutate(Region = factor(Region, levels=unique(Region)))

# Cluster prop estimates ----------------------------------------------
set.seed(seed)
cluster_prop_meta_by_region <- meta::metaprop(
    n_isolates_clusters, n_isolates, studlab=site,
    data=all_sites_clusters_summary, 
    subgroup=Region, method = "GLMM", method.ci = "CP", # method.tau="DL",
    sm="PLOGIT")
# forest plot
cluster_prop_meta_by_region_forest <- {
    meta::forest(cluster_prop_meta_by_region, xlab="Cluster proportion", common=F, 
                 sortvar=TE, print.I2.ci=T,
                 leftcols = c("studlab","event","n","effect.ci"), rightcols=F,
                 leftlabs = c("Site","Cluster\nisolates","Total\nisolates","Proportion\n[95% CI]"),
                 colgap.left=unit(0.7, "cm"), plotwidth="10cm",
                 col.square="#8b0000", col.square.lines="#8b0000", 
                 col.diamond.random="#ffc1c1", 
                 xlim=c(0,1), fontsize=15, digits=2);
    grid.grab()
}

# Transmission prop estimates ----------------------------------------------
set.seed(seed)
transmi_prop_meta_by_region <- meta::metaprop(
    n_transmission_isolates, n_isolates, studlab=site,
    data=all_sites_clusters_summary, 
    subgroup=Region, method = "GLMM", method.ci = "CP", # method.tau="DL",
    sm="PLOGIT")
# forest plot
transmi_prop_meta_by_region_forest <- {
    meta::forest(transmi_prop_meta_by_region, xlab="Transmission proportion", common=F, 
                 sortvar=TE, print.I2.ci=T,
                 leftcols = c("studlab","event","n","effect.ci"), rightcols=F,
                 leftlabs = c("Site","Transmitted\nisolates","Total\nisolates","Proportion\n[95% CI]"),
                 colgap.left=unit(0.7, "cm"), plotwidth="10cm",
                 col.square="navyblue", col.square.lines="navyblue",
                 col.diamond.random="lightblue", 
                 xlim=c(0,1), fontsize=15, digits=2);
    grid.grab()
}

get_metaprop_subgroup_estimates(transmi_prop_meta_by_region)

# ggsave(file=file.path(figs_dir, "FigureS7a_PooledEstimatesRegion.png"), 
#        plot=ggarrange(cluster_prop_meta_by_region_forest, labels=c("(A) Cluster proportion"),
#                       font.label = list(size = 18)), 
#        width=11, height=11, bg='white')
# ggsave(file=file.path(figs_dir, "FigureS7a_PooledEstimatesRegion.tif"), 
#        plot=ggarrange(cluster_prop_meta_by_region_forest, labels=c("(A) Cluster proportion"),
#                       font.label = list(size = 18)), 
#        width=11, height=11, bg='white')
# ggsave(file=file.path(figs_dir, "FigureS7b_PooledEstimatesRegion.png"), 
#        plot=ggarrange(transmi_prop_meta_by_region_forest, labels=c("(B) Transmission proportion"),
#                       font.label = list(size = 18)), 
#        width=11, height=11, bg='white')
# ggsave(file=file.path(figs_dir, "FigureS7b_PooledEstimatesRegion.tif"), 
#        plot=ggarrange(transmi_prop_meta_by_region_forest, labels=c("(B) Transmission proportion"),
#                       font.label = list(size = 18)), 
#        width=11, height=11, bg='white')

```

<br><br>

```{r facilities_characteristics, warning=F}
# data
facilities_df <- read_tsv("./data/facilityInfo.tsv", show_col_types=F) %>% 
    mutate(Site = factor(Site, levels = unique(Site))) %>% 
    select(Site, Region, Country, Location, 
           Facility_type = `Facility type`, 
           Facility_size = `Facility size`,
           Piped_water = `Piped water available`,
           Neonatal_beds = `No. neonatal beds`,
           Onsite_surgery = `Onsite neonatal surgical facilities`) %>% 
    separate_wider_regex(Facility_size, patterns = c(Facility_size = ".+", ":.+")) %>% 
    mutate(Facility_type = gsub(" Hospital$", "", Facility_type)) %>% 
    mutate(Piped_water = case_when(
        Piped_water == "Occassionally" ~ "Sometimes",
        Piped_water == "Most of the time" ~ "Most times",
        TRUE ~ Piped_water)) %>%
    arrange(Neonatal_beds) %>% 
    mutate(Neonatal_beds_grouped = case_when(
        Neonatal_beds > 100 ~ ">100",
        Neonatal_beds >= 50 ~ "50-100",
        Neonatal_beds < 50 ~ "<50",
        TRUE ~ NA_character_
    )) %>% 
    mutate(Neonatal_beds_grouped = factor(Neonatal_beds_grouped, levels=unique(Neonatal_beds_grouped))) %>% 
    mutate(
        Location = factor(Location, levels = c("Rural", "Urban")),
        Facility_type = factor(Facility_type, levels = c("District", "Regional", "Tertiary")),
        Facility_size = factor(Facility_size, levels = c("Small", "Medium", "Large", "Very Large")),
        Piped_water = factor(Piped_water, levels = c("Sometimes", "Most times", "Always")),
        Onsite_surgery = factor(Onsite_surgery, levels = c("No", "Yes"))) %>% 
    arrange(Site)
write_tsv(facilities_df %>% 
              select(Region, Country, Location, Facility_size, # Facility_type,
                     Neonatal_beds, Piped_water, Onsite_surgery, Site), 
          file.path(tables_dir, "Table2_FacilityInfo.tsv"))

facilities_df_and_transmission_data <- all_sites_clusters_summary %>% 
    arrange(desc(cluster_prop)) %>% 
    left_join(facilities_df %>% select(-c(Region, Country)), 
              by = c("site"="Site")) %>% 
    filter(!is.na(Facility_type))

# Heatmap of site characteristics vs cluster prop
facs_heatmap_c <- plot_facility_data_heatmap(facilities_df, cluster_prop_meta_df)


# Combine forest plots with site characteristics heatmap
clust_meta_with_facility_heatmap <- (
        cluster_prop_meta_plots$p.left |  
        cluster_prop_meta_plots$p.right + theme(legend.key.height = unit(2, "mm")) | 
        facs_heatmap_c$p1 + theme(axis.text.x = element_text(size=8)) | 
        facs_heatmap_c$p3 + theme(axis.text.x = element_text(size=8)) | 
        facs_heatmap_c$p4 + theme(axis.text.x = element_text(size=8)) | 
        facs_heatmap_c$p5 + theme(axis.text.x = element_text(size=8), 
                                  legend.key.height = unit(3, "mm")) | 
        facs_heatmap_c$p6 + theme(axis.text.x = element_text(size=8))
    ) + 
    plot_annotation(
        title = "(A)",
        theme = theme(plot.title = element_text(size=9, face="bold"), plot.title.position = "panel")) +
    plot_layout(widths=c(0.7,0.7,0.04,0.04,0.04,0.04,0.04), guides="collect") &
    theme(plot.tag = element_text(face="bold", vjust=-3.3),
          legend.text = element_text(size=8), legend.title = element_text(size=10),
          legend.position = "bottom", legend.direction = "vertical", legend.box = "horizontal")

ggsave(file=file.path(figs_dir, "FigureS5a_PooledEstimatesWithFacilityDetails.png"),
       plot=clust_meta_with_facility_heatmap, width=8, height=6.5, bg='white')
ggsave(file=file.path(figs_dir, "FigureS5a_PooledEstimatesWithFacilityDetails.tif"),
       plot=clust_meta_with_facility_heatmap, width=8, height=6.5, bg='white')


```

#### Pooled estimates by site characteristics
```{r facilities_metareg}
#########################################################
# Meta regression
#########################################################
covariates = c("Onsite_surgery", "Facility_size", "Neonatal_beds", "Piped_water")
init_mt <- meta::metaprop(
        n_isolates_clusters, n_isolates, studlab=site, 
        data=facilities_df_and_transmission_data, 
        method="GLMM", method.ci="CP", sm="PLOGIT", common=F)
meta_by_facs_xters <- metareg(init_mt, paste(covariates, collapse = " + "))
meta_by_facs_xters_df <- parse_metareg_model(meta_by_facs_xters)
summary(meta_by_facs_xters)
# meta_by_facs_xters_df %>% 
#     select(Predictor, OR, lower, upper, P) %>% 
#     mutate(OR = sprintf("%.2f", OR), `95% CI` = sprintf("%.2f - %.2f", lower, upper)) %>% 
#     select(Predictor, `Odds ratio` = OR, `95% CI`, P) %>% 
#     write_tsv(., file.path(tables_dir, "TableSX_FacilitiesVsTransmissionReg.tsv"), na="-")


#########################################################
# Fishers exact tests 
#########################################################
fac_vars <- c("Onsite_surgery", "Facility_size", "Neonatal_beds_grouped", "Piped_water")
fac_vars_transmission_tests <- purrr::map_dfr(fac_vars, function(v) {
    summary_df <- facilities_df_and_transmission_data %>%
        group_by(Group = !!sym(v)) %>%
        reframe(f = sum(n_isolates_clusters), n = sum(n_isolates), prop = f/n)
    pairs <- combn(nrow(summary_df), 2) # unique pairs in each covariate
    # run Fisher test on each pair
    map_df(1:ncol(pairs), function(i) {
        idx <- pairs[, i]
        g1 <- summary_df[idx[1], ]; g2 <- summary_df[idx[2], ]
        mat <- matrix(c(g1$f, g1$n - g1$f, g2$f, g2$n - g2$f), nrow = 2, byrow = TRUE)
        data.frame(variable = v,
                   comparison = paste(g1$Group, "vs", g2$Group),
                   pct_A = round(g1$prop * 100, 1),
                   pct_B = round(g2$prop * 100, 1),
                   p.value = fisher.test(mat)$p.value)
        })
    }) %>%
    mutate(padj = p.adjust(p.value, method = "bonferroni")) %>% 
    mutate(signif = padj < 0.05) %>% 
    mutate(padj_report = case_when(
      padj < 0.001 ~ "p<0.001",
      padj < 0.01  ~ sprintf("p=%.4f", padj),
      padj < 0.05  ~ sprintf("p=%.2f", padj),
      TRUE      ~ "ns"
    ))
# fac_vars_transmission_tests %>% select(-padj_report) %>% 
#     write_tsv(., file.path(tables_dir, "TableSX_FacilitiesVsTransmission.tsv"), na="-")

#########################################################
# Plots
#########################################################
# Meta regression plot
meta_by_facs_xters_plot <- meta_by_facs_xters_df %>%  
    mutate(Group = case_when(
        grepl("^Onsite_surgery", Predictor) ~ "Onsite\nsurgery",
        grepl("^Facility_size", Predictor) ~ "Facility\nsize",
        grepl("^Piped_water", Predictor) ~ "Piped\nwater",
        grepl("^Neonatal_beds", Predictor) ~ "Neonatal\nbeds",
        TRUE ~ NA_character_
    )) %>% 
    mutate(Group = factor(Group, levels = c("Onsite\nsurgery", "Facility\nsize", 
                                            "Neonatal\nbeds", "Piped\nwater"))) %>% 
    mutate(Predictor = gsub("Onsite_surgery|Facility_size|Piped_water", "", Predictor)) %>% 
    mutate(Predictor = if_else(Predictor == "Neonatal_beds", "No. of beds", Predictor)) %>% 
    mutate(signif = case_when(
        P < 0.05 ~ "P < 0.05",
        P >= 0.05 ~ "P ≥ 0.05",
        TRUE ~ NA_character_
    )) %>% 
    arrange(OR) %>% mutate(Predictor = factor(Predictor, levels = unique(Predictor))) %>% 
    ggplot(aes(y = Predictor)) +
    geom_pointrange(aes(x=OR, xmin=lower, xmax=upper, color=signif),
                    position = position_dodge(width=.6), shape=16, size=.85) + 
    scale_color_manual(values = c("darkred", "#FF8A8A"), name=NULL) +
    geom_vline(xintercept = 1, lty = 2) +
    labs(x = "Odds ratio with 95% CI", title="Association of facility characteristics\nwith transmission") +
    facet_grid(rows = vars(Group), scales = "free_y", space = "free_y") +
    theme_bw() +
    theme(panel.grid.minor = element_blank(), axis.text = element_text(size=12),
          axis.title = element_text(size=14), strip.text = element_text(size=12),
          legend.text = element_text(size=12), panel.grid.major.y = element_blank(),
          legend.position = "none",
          legend.justification = "top", legend.key.height = unit(24, "pt"))

meta_by_facs_xters_plot

# Boxplots of site characteristics vs cluster proportion
facs_clust_prop_boxplots_lst <- plot_clust_prop_by_facility_data(
    facilities_df_and_transmission_data %>% rename("No. neonatal beds" = "Neonatal_beds_grouped"),
    c("Onsite_surgery", "Facility_size", "Piped_water", "No. neonatal beds"))

for (i in c(1:length(facs_clust_prop_boxplots_lst))) { 
    facs_clust_prop_boxplots_lst[[i]] <- facs_clust_prop_boxplots_lst[[i]] +
        theme(axis.title.y=element_blank(), axis.text.y=element_blank(),
              axis.text.x = element_text(angle=45, hjust=1),
              plot.title = element_text(hjust = 0.5))
}


# Combine plots #################################

# combined_facs_clust_prop_plots <-
#     wrap_elements(
#         meta_by_facs_xters_plot +
#         plot_annotation(
#             title = "(B)", theme = theme(plot.title = element_text(size=14, face="bold"), 
#                                          plot.title.position = "panel"))
#     ) + 
#     wrap_elements((
#         facs_clust_prop_boxplots_lst[[1]] + 
#             theme(axis.title.y=element_text(size=12, angle=90, vjust=1), 
#                   axis.text.y=element_text(size=12)) +
#         facs_clust_prop_boxplots_lst[[2]] +
#         facs_clust_prop_boxplots_lst[[3]] + 
#             theme(axis.title.y=element_text(size=12, angle=90, vjust=1), 
#                   axis.text.y=element_text(size=12)) +
#         facs_clust_prop_boxplots_lst[[4]]) + 
#         plot_annotation(title = "(C)", tag_levels="i", tag_prefix="(", tag_suffix=")",
#                         theme = theme(plot.title = element_text(size=14, face="bold"), 
#                                              plot.title.position = "panel"))
#     ) + 
#     plot_layout(nrow=1, widths = c(0.4,0.6)) &
#     theme(plot.tag = element_text(size=14)) 

combined_facs_clust_prop_plots <-
    (
        facs_clust_prop_boxplots_lst[[1]] + 
            theme(axis.title.y=element_text(size=12, angle=90, vjust=1), 
                  axis.text.y=element_text(size=12)) +
        facs_clust_prop_boxplots_lst[[2]] +
        facs_clust_prop_boxplots_lst[[3]] + 
            theme(axis.title.y=element_text(size=12, angle=90, vjust=1), 
                  axis.text.y=element_text(size=12)) +
        facs_clust_prop_boxplots_lst[[4]]
    ) + 
    plot_annotation(title = "(B)", tag_levels="i", tag_prefix="(", tag_suffix=")",
                    theme = theme(plot.title = element_text(size=14, face="bold"), 
                                         plot.title.position = "panel")) +
    plot_layout(nrow=1) &
    theme(plot.tag = element_text(size=14)) 

ggsave(file=file.path(figs_dir, "FigureS5b_ClusterPropByFacilityCharacteristics.png"),
       plot=combined_facs_clust_prop_plots, width=14, height=5, bg='white')
ggsave(file=file.path(figs_dir, "FigureS5b_ClusterPropByFacilityCharacteristics.tif"),
       plot=combined_facs_clust_prop_plots, width=14, height=5, bg='white')


#########################################################
# Median cluster prop within facs xters categories
#########################################################
facilities_df_and_transmission_data %>%
    mutate(cluster_prop = round(cluster_prop*100, 1)) %>%
    group_by(Piped_water) %>%
    reframe(n = n(), 
            Cluster_prop = paste0(
                median(cluster_prop, na.rm=T), "% [",
                paste0(range(cluster_prop), collapse = "-"), "%]"
            )
    )
facilities_df_and_transmission_data %>%
    mutate(cluster_prop = round(cluster_prop*100, 1)) %>%
    group_by(Onsite_surgery) %>%
    reframe(n = n(), 
            Cluster_prop = paste0(
                median(cluster_prop, na.rm=T), "% [",
                paste0(range(cluster_prop), collapse = "-"), "%]"
            )
    )
facilities_df_and_transmission_data %>%
    mutate(cluster_prop = round(cluster_prop*100, 1)) %>%
    group_by(Facility_size) %>%
    reframe(n = n(), 
            Cluster_prop = paste0(
                median(cluster_prop, na.rm=T), "% [",
                paste0(range(cluster_prop), collapse = "-"), "%]"
            )
    )


#########################################################
# Correlation
#########################################################
corr_clust_vs_n_cases <- ggplot(data = facilities_df_and_transmission_data, 
                                aes(x = cluster_prop, y = n_isolates)) +
    ggpmisc::stat_correlation(ggpmisc::use_label(c("R2", "P"), method="pearson"), small.r=T) +
    ggpmisc::stat_poly_line(color = "black") +
    geom_point() +
    scale_x_continuous(breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
    labs(x = "Cluster proportion", y = "Number of cases",
         title = "Correlation between cluster proportion and number of cases") +
    theme_bw() +
    theme(axis.text = element_text(size = 10), axis.title = element_text(size = 12),
          legend.text = element_text(size = 10), 
          plot.title = element_text(size = 12), legend.position = "bottom")
corr_clust_vs_n_cases

corr_cases_vs_beds <- ggplot(data = facilities_df_and_transmission_data, 
                                aes(x = n_isolates, y = Neonatal_beds)) +
    ggpmisc::stat_correlation(ggpmisc::use_label(c("R2", "P"), method="pearson"), small.r=T) +
    ggpmisc::stat_poly_line(color = "black") +
    geom_point() +
    labs(x = "Number of cases", y = "No. neonatal beds",
         title = "Correlation between no. of cases and no. of neonatal beds") +
    theme_bw() +
    theme(axis.text = element_text(size = 10), axis.title = element_text(size = 12),
          legend.text = element_text(size = 10), 
          plot.title = element_text(size = 12), legend.position = "bottom")
corr_cases_vs_beds
```

<br><br>  

### Transmission clones  

```{r clones_summary}
# CHeck clusters with dissimilar STs (LVs)
# These will have NA in some cluster summary columns as they dont have unique STs
# So rename these LV STs (only those specifically forming clusters with the ST) to the predominant ST
dissim_ST_clust <- epi_snp_clusters %>% filter(!is.na(Cluster)) %>% 
    group_by(Cluster) %>% count(ST) %>% select(-n)
dissim_ST_clust <- bind_rows(dissim_ST_clust[duplicated(dissim_ST_clust$Cluster),], 
          dissim_ST_clust[duplicated(dissim_ST_clust$Cluster, fromLast = T),]) 
dissim_ST_clust %<>% 
    arrange(Cluster, ST) %>% 
    group_by(Cluster) %>% 
    summarise(STs = paste0(ST, collapse = "; ")) %>% 
    rowwise() %>% 
    # group under predominant ST. First following arrnage
    mutate(ST_pred = case_when(
        grepl(";", STs) ~ str_split(STs, "; ", simplify=T)[1], 
        TRUE ~ STs)) %>% 
    ungroup() %>% # select(Cluster, ST_pred) %>% 
    right_join(dissim_ST_clust, by = "Cluster")

epi_snp_clusters_wt_upd_STS <- epi_snp_clusters %>% 
    left_join(dissim_ST_clust, by = c("Cluster", "ST")) %>% 
    mutate(ST = if_else(is.na(ST_pred), ST, ST_pred)) %>% 
    mutate(ST_annot = if_else(is.na(ST_pred), ST, paste0(ST_pred, "*"))) %>% 
    select(-ST_pred)

all_clusters_info_wt_upd_STs <- all_clusters_info %>% 
    # rename dissim STs in clusters to their predom STs
    left_join(distinct(dissim_ST_clust, Cluster, STs, ST_pred),
              by = c("Cluster"="Cluster", "ST"="STs")) %>% 
    mutate(ST = if_else(is.na(ST_pred), ST, ST_pred)) %>% select(-ST_pred)


# Dereplicated clusters data (one row per cluster; one row per singleton)
epi_snp_clusters_dereplicated <- epi_snp_clusters_wt_upd_STS %>% 
    # Deduplicate clusters
    filter(!is.na(Cluster)) %>% arrange(Cluster, Year) %>% 
    group_by(Cluster) %>% 
    # Consensus (mode) KO for the cluster
    mutate(K_locus = modeest::mfv1(K_locus), 
           O_group = modeest::mfv1(O_group)) %>% 
    # Consensus res score for the cluster (higher with tie)
    mutate(resistance_score = modeest::mfv1(resistance_score)) %>% 
    slice_min(Year, with_ties = F) %>% ungroup() %>% 
    # Add in non-cluster isolates
    bind_rows(epi_snp_clusters_wt_upd_STS %>% filter(is.na(Cluster)))


clones_summary <- epi_snp_clusters_wt_upd_STS %>% 
    group_by(ST) %>% 
    reframe(
        n_clusters = n_distinct(Cluster, na.rm = T),
        n_isolates = n(),
        n_isolates_clusters = n_distinct(id[!is.na(Cluster)]),
        n_isolates_singletons = n_distinct(id[is.na(Cluster)]),
        n_countries = n_distinct(Country, na.rm=T),
        n_countries_clusters = n_distinct(Country[!is.na(Cluster)], na.rm=T),
        n_countries_singletons = n_distinct(Country[is.na(Cluster)], na.rm=T),
        n_sites = n_distinct(Site, na.rm=T),
        n_sites_clusters = n_distinct(Site[!is.na(Cluster)], na.rm=T),
        n_sites_singletons = n_distinct(Site[is.na(Cluster)], na.rm=T),
        n_clones = n_distinct(ST, na.rm=T),
        n_clones_clusters = n_distinct(ST[!is.na(Cluster)], na.rm=T),
        n_clones_singletons = n_distinct(ST[is.na(Cluster)], na.rm=T),
        mean_res_score = round(mean(resistance_score, na.rm=T),1),
        mean_res_score_clusters = round(mean(resistance_score[!is.na(Cluster)], na.rm=T),1),
        mean_res_score_singletons = round(mean(resistance_score[is.na(Cluster)], na.rm=T),1),
    ) %>% 
    arrange(desc(n_clusters)) %>% 
    left_join(
         all_clusters_info_wt_upd_STs %>%
            group_by(ST) %>% 
            reframe(median_cluster_size = median(`N isolates`, na.rm=T)),
        by='ST')

caption_text <- glue("<p style='text-align:left; padding-top:16px'><b>Clones summary </b></p>")
caption_html <- htmltools::HTML(caption_text)
DT::datatable(clones_summary,
              caption = knitr::asis_output(caption_html),
              options = list(autoWidth = TRUE, pageLength = 15, 
                             scrollX = TRUE, scrollY = TRUE, rownames = FALSE)
              )

# Common transmission clones (Occurring in ≥3 clusters and ≥2 sites)
top_clones_summary <- clones_summary %>%
    filter(n_clusters >= 3 & n_sites_clusters >= 2) %>%
    arrange(desc(n_isolates_clusters))

# All clusters summary
all_clusters_summary <- summarise_cluster_nodates(epi_snp_clusters_wt_upd_STS)
DT::datatable(all_clusters_summary, caption="Clusters summary (with raw overall estimates)")
```

<br><br>

```{r combine_cluster_plots, fig.width=12, fig.height=15}
###
# REQUIRES DATES SO CODE BELOW WILL NOT WORK ON PUBLISHED DATA WITHOUT DATES
###


# all_cluster_plots_lst <- lapply(studies, function(x) {
#     d <- epi_snp_clusters_wt_upd_STS %>% filter(Study == x) %>% 
#         mutate(ST = ST_annot)
#     study_lab <- unique(d$Study_lab)
#     plot_clusters2(d, color_column="ST", dodge_height=0.2) + 
#         guides(fill="none", color='none', size=guide_legend(nrow=1)) + 
#         labs(title=study_lab, y="ST") +
#         theme(plot.title=element_text(size=14), legend.position = "top", 
#               legend.justification="left", legend.margin=margin(-5, 0, -7, 0), 
#               legend.spacing.y = unit(0, "mm"), axis.title = element_blank(),
#               axis.text = element_text(size=12), legend.text = element_text(size=12),
#               legend.title = element_text(size=12)) +
#         geom_line(aes(group = Cluster, alpha = 0.25), color = "black", linewidth=1,
#                   position=ggstance::position_dodgev(height = 0.2))
# })
# all_cluster_plots <- ggarrange(plotlist=all_cluster_plots_lst, common.legend=F, nrow=4, ncol=3)
# all_cluster_plots <- annotate_figure(
#     all_cluster_plots, 
#     left = text_grob("Sequence type", vjust = 1, size=16, rot=90),
#     bottom = text_grob("Isolation date", size=16)
# )
# all_cluster_plots
# ggsave(file=file.path(figs_dir, "FigureS3_ClustersPerSite.png"), 
#        plot=all_cluster_plots, width=15, height=16, bg="white")
# ggsave(file=file.path(figs_dir, "FigureS3_ClustersPerSite.tif"), 
#        plot=all_cluster_plots, width=15, height=16, bg="white")


```

<br>

```{r ko_summary, warning=F}
k_summary <- summarise_ko_by_case_type(epi_snp_clusters_dereplicated, "K_locus")

o_summary <- summarise_ko_by_case_type(epi_snp_clusters_dereplicated, "O_group")


DT::datatable(k_summary, caption="KL summary",
              options=list(autoWidth=TRUE, pageLength=15, scrollX=TRUE, 
                           scrollY=TRUE, rownames=FALSE))
DT::datatable(o_summary, caption="O summary",
              options=list(autoWidth=TRUE, pageLength=15, scrollX=TRUE, 
                           scrollY=TRUE, rownames=FALSE))

# write
write_tsv(o_summary, file.path(tables_dir, "TableS9_OSummary.tsv"))
write_tsv(k_summary, file.path(tables_dir, "TableS10_KLSummary.tsv"))


print(paste(sum(!is.na(k_summary$Clusters)), "distinct K loci were found associated with transmission clusters.", sum(!is.na(k_summary$Clusters) & !is.na(k_summary$Singletons)), "of these were also detected in singletons"))
k_summary %>% filter(!is.na(k_summary$Clusters) & is.na(k_summary$Singletons)) %>%
    pull(K_locus) %>% paste0(collapse = ", ")

# Cluster sizes for KLs associated with clusters
epi_snp_clusters %>% 
    filter(!is.na(Cluster)) %>% group_by(K_locus) %>% 
    count(Cluster) %>% reframe(mdn = median(n), rng = paste0(range(n), collapse = " - ")) %>%
    mutate(cluster_size = glue::glue("{mdn} [{rng}]")) %>% select(-c(mdn,rng))

# Most common K loci associated with transmission
# Defined as present in transmitted isolates in ≥3 clusters and ≥2 sites
top_k_summary <- k_summary %>% 
    separate(Clusters, into = "n_clusters", sep=" ", remove=F, extra="drop") %>% 
    mutate(n_clusters = as.numeric(n_clusters)) %>% 
    filter(n_clusters >= 3 & n_sites_clusters >= 2) %>%
    arrange(desc(n_clusters))

# top_k_summary %>% arrange(desc(n_clusters)) %>% 
#     mutate(rep = glue("{K_locus} ({n_clusters} clusters in {n_sites_clusters} sites)")) %>%
#     pull(rep) %>% paste0(collapse = ", ")


paste0(top_k_summary %>% filter(n_sites_clusters/n_sites*100 >= 33.3) %>% nrow(),
       " KL associated with transmission clusters in at least a 3rd of sites"
)

# Report clone association of the top K loci associated with transmission
epi_snp_clusters_wt_upd_STS %>% filter(K_locus == "KL102") %>% 
    mutate(case_type = if_else(is.na(Cluster), "Single", "Cluster")) %>% 
    count(case_type, ST) %>% arrange(case_type) %>% 
    reframe(sum(n), .by = c(case_type, ST)) %>% 
    filter(case_type == "Cluster")
epi_snp_clusters_wt_upd_STS %>% filter(K_locus == "KL102" & !is.na(Cluster)) %>% 
    group_by(Cluster) %>% reframe(Clone = paste0(unique(ST), collapse = "; ")) %>% 
    count(Clone, name = "# KL102 clusters")

epi_snp_clusters_wt_upd_STS %>% filter(K_locus == "KL25") %>% 
    mutate(case_type = if_else(is.na(Cluster), "Single", "Cluster")) %>% 
    count(case_type, ST) %>% arrange(case_type) %>% 
    reframe(sum(n), .by = c(case_type, ST)) %>% 
    filter(case_type == "Cluster")
```



```{r transmission_predictors, warning=F}
# Regression model to identify factors contributing to transmission risk
# common_ST_df <- epi_snp_clusters_dereplicated %>% 
#     group_by(ST) %>% count() %>% ungroup() %>% 
#     mutate(ST_prev_grp = if_else(n>2, "Common_ST2", "Rare_ST")) 
# common_ST2 <- common_ST_df %>% 
#     filter(ST_prev_grp == "Common_ST2") %>% 
#     arrange(desc(n)) %>% pull(ST)
# common_ST_df %>% group_by(ST_prev_grp) %>% 
#     reframe(n = sum(n)) %>% 
#     mutate(pct = round(n/sum(n)*100,1))

common_KL_df <- epi_snp_clusters_dereplicated %>% 
    group_by(K_locus) %>% count() %>% ungroup() %>% 
    mutate(KL_prev_grp = if_else(n > (nrow(epi_snp_clusters_dereplicated)*0.01), 
                                 "Common_KL", "Rare_KL"))
common_KL <- common_KL_df %>% 
    filter(KL_prev_grp == "Common_KL") %>% 
    arrange(desc(n)) %>% pull(K_locus)
# common_KL_df %>% group_by(KL_prev_grp) %>% 
#     reframe(n = sum(n)) %>% 
#     mutate(pct = round(n/sum(n)*100,1))
common_KL_df %>% 
    mutate(grp = if_else(K_locus %in% common_KL[1:10], "Top 10", "Others")) %>% 
    group_by(grp) %>%
    reframe(n = sum(n)) %>%
    mutate(pct = round(n/sum(n)*100,1))

common_STs_resistance <- epi_snp_clusters_dereplicated %>% 
    mutate(common_ST = if_else(ST %in% top_clones_summary$ST, ST, "Other STs")) %>% 
    mutate(case_type = if_else(is.na(Cluster), "Single", "Cluster")) %>%
    mutate(ESBL = if_else(resistance_score >= 1, "+", "-")) %>%
    mutate(CP = if_else(resistance_score >= 2, "+", "-")) %>%
    bind_rows(., mutate(., case_type = "All")) %>% 
    group_by(common_ST, case_type) %>%
    reframe(N = n(),
            nESBL = n_distinct(id[ESBL == "+"]), pctESBL = round(nESBL/n()*100,1),
            nCP = n_distinct(id[CP == "+"]), pctCP = round(nCP/n()*100,1)) %>% 
    mutate(ESBL = sprintf("%.0f (%.1f%%)", nESBL, pctESBL),
           CP = sprintf("%.0f (%.1f%%)", nCP, pctCP),
           N = as.character(N)) %>% 
    select(-c(nESBL, pctESBL, nCP, pctCP)) %>% 
    pivot_longer(-c(common_ST, case_type), names_to = "names", values_to = "values") %>% 
    mutate(grp = glue("{case_type}: {names}")) %>% 
    pivot_wider(id_cols = c(common_ST), names_from = grp, values_from = values, values_fill = "-") %>% 
    mutate(common_ST = factor(common_ST, levels = c(top_clones_summary$ST, "Other STs"))) %>% 
    arrange(common_ST)

# Model data
trs_pred_data <- epi_snp_clusters_dereplicated %>% 
    select(id, Site, Cluster, ST, K_locus, O_type, resistance_score) %>% 
    mutate(Site = factor(Site, levels = sites)) %>% 
    mutate(Case = if_else(is.na(Cluster), 0, 1)) %>%
    mutate(commonST_ = if_else(ST %in% top_clones_summary$ST, ST, "Other STs")) %>% 
    mutate(commonST_ = factor(commonST_, levels = c("Other STs", top_clones_summary$ST))) %>% 
    mutate(ESBL = if_else(resistance_score >= 1, "+", "-"),
           ESBL = factor(ESBL, levels = c("-", "+"))) %>%
    mutate(CP = if_else(resistance_score >= 2, "+", "-"),
           CP = factor(CP, levels = c("-", "+"))) %>%
    mutate(O_type_grp = case_when(
        O_type %in% c("O12", "O13", "unknown") ~ "Other",
        grepl("^O1a", O_type) ~ "O1",
        grepl("^O2a", O_type) ~ "O2",
        grepl("^O3", O_type) ~ "O3",
        grepl("^O4", O_type) ~ "O4",
        grepl("^O5", O_type) ~ "O5",
        TRUE ~ NA_character_
    )) %>% 
    mutate(O_type_grp = factor(O_type_grp, levels=c("O1","O2","O3","O4","O5","Other"))) %>% 
    mutate(commonKL_ = if_else(K_locus %in% common_KL[1:10], K_locus, "Other KLs")) %>% 
    mutate(commonKL_ = factor(commonKL_, levels = c("Other KLs", common_KL[1:10]))) %>% 
    mutate(K_locus = factor(K_locus, levels=k_summary$K_locus))

############################################
# Fit models
###########################################
# AMR
trs_model_Amr <- logistf::logistf("Case ~ ESBL + CP", data=trs_pred_data)
trs_model_Amr_df <- parse_logistf_model(trs_model_Amr)

# AMR + Site
trs_model_AmrSite <- logistf::logistf("Case ~ ESBL + CP + Site", data=trs_pred_data)
trs_model_AmrSite_df <- parse_logistf_model(trs_model_AmrSite) %>% 
    mutate(Predictor = gsub("SiteSite", "Site", Predictor))

# AMR + Site + ST
trs_model_AmrSiteSt <- logistf::logistf("Case ~ ESBL + CP + Site + commonST_", data=trs_pred_data)
trs_model_AmrSiteSt_df <- parse_logistf_model(trs_model_AmrSiteSt) %>% 
    mutate(Predictor = gsub("^commonST_", "", Predictor)) %>% 
    mutate(Predictor = gsub("SiteSite", "Site", Predictor))

# AMR + ST
trs_model_AmrSt <- logistf::logistf("Case ~ ESBL + CP + commonST_", data=trs_pred_data)
trs_model_AmrSt_df <- parse_logistf_model(trs_model_AmrSt) %>% 
    mutate(Predictor = gsub("^commonST_", "", Predictor))

# ST
trs_model_St <- logistf::logistf("Case ~ commonST_", data=trs_pred_data)
trs_model_St_df <- parse_logistf_model(trs_model_St) %>% 
    mutate(Predictor = gsub("^commonST_", "", Predictor))

# ST + Site
trs_model_StSite <- logistf::logistf("Case ~ commonST_ + Site", data=trs_pred_data)
trs_model_StSite_df <- parse_logistf_model(trs_model_StSite) %>% 
    mutate(Predictor = gsub("^commonST_", "", Predictor)) %>% 
    mutate(Predictor = gsub("SiteSite", "Site", Predictor))


###########################################
# compare models
###########################################
anova(trs_model_Amr, trs_model_AmrSite) # Does adding site improve model with AMR
anova(trs_model_AmrSite, trs_model_AmrSiteSt) # Does adding ST improve model with AMR + Site
anova(trs_model_Amr, trs_model_AmrSt) # Does adding ST alone improve AMR model ?
anova(trs_model_St, trs_model_StSite) # Does adding site improve ST model ?
anova(trs_model_St, trs_model_AmrSiteSt) # Does adding site and amr improve ST model?
anova(trs_model_StSite, trs_model_AmrSiteSt) # Does adding amr improve ST + Site model?


###########################################
# Best ST model for transmission
###########################################
final_trs_model <- trs_model_AmrSiteSt
final_trs_model_df <- trs_model_AmrSiteSt_df


############################################
# Test models when KL added
trs_model_AmrK <- logistf::logistf("Case ~ ESBL + CP + commonKL_",
                                         data=trs_pred_data)
trs_model_AmrSiteK <- logistf::logistf("Case ~ ESBL + CP + Site + commonKL_",
                                         data=trs_pred_data)
trs_model_AmrStK <- logistf::logistf("Case ~ ESBL + CP + commonST_ + commonKL_",
                                         data=trs_pred_data)
trs_model_AmrSiteStK <- logistf::logistf("Case ~ ESBL + CP + Site + commonST_ + commonKL_",
                                         data=trs_pred_data)

anova(trs_model_Amr, trs_model_AmrK) # Does adding KL improve model with AMR
anova(trs_model_AmrSiteSt, trs_model_AmrSiteStK) # Does adding KL improve model with AMR + ST + Site
# including KL in the model with ESBL and carbapenemase, with or without ST and/or Site, 
# did not improve model fit (p>0.38) and no KLs were significant?

trs_model_AmrSiteStK_df <- parse_logistf_model(trs_model_AmrSiteStK) %>% 
    mutate(Predictor = gsub("^commonST_", "", Predictor)) %>% 
    mutate(Predictor = gsub("SiteSite", "Site", Predictor)) %>% 
    mutate(Predictor = gsub("^commonKL_", "", Predictor))

# KL62 singificant
epi_snp_clusters_dereplicated %>% 
    filter(K_locus == "KL62")  %>% 
    left_join(all_clusters_info %>% select(Cluster, `N isolates`), by="Cluster") %>% 
    reframe(
        Intros = n(), 
        Clusters = paste0(n_distinct(Cluster, na.rm=T), ": ", sum(`N isolates`, na.rm=T)),
        STs = paste0("n=", n_distinct(ST), ": ", paste0(unique(sort(ST)), collapse = ",")), 
        Sites = paste0("n=", n_distinct(Site), ": ", paste0(unique(sort(Site)), collapse = ","))
    ) %>% 
    view()

############################################
# Test models when O types added
trs_model_AmrO <- logistf::logistf("Case ~ ESBL + CP + O_type_grp",
                                   data=trs_pred_data)
trs_model_AmrSiteO <- logistf::logistf("Case ~ ESBL + CP + Site + O_type_grp",
                                       data=trs_pred_data)
trs_model_AmrStO <- logistf::logistf("Case ~ ESBL + CP + commonST_ + O_type_grp",
                                     data=trs_pred_data)
trs_model_AmrSiteStO <- logistf::logistf("Case ~ ESBL + CP + Site + commonST_ + O_type_grp",
                                         data=trs_pred_data)

anova(trs_model_Amr, trs_model_AmrO) # Does adding O types improve model with AMR
anova(trs_model_AmrSiteSt, trs_model_AmrSiteStO) # Does adding O types improve model with AMR + ST + Site
# including O types in the model with ESBL and carbapenemase, with or without ST and Site, 
# did not improve model fit (p>0.2) and no O types were significant?

trs_model_AmrSiteStO_df <- parse_logistf_model(trs_model_AmrSiteStO) %>% 
    mutate(Predictor = gsub("^commonST_", "", Predictor)) %>% 
    mutate(Predictor = gsub("SiteSite", "Site", Predictor)) %>% 
    mutate(Predictor = gsub("^O_type_grp", "", Predictor))


############################################
# Plot best ST model
############################################
trs_model_plot <-  final_trs_model_df %>%  
    mutate(cat = case_when(
        Predictor %in% c("ESBL+", "CP+") ~ "AMR",
        grepl("^ST", Predictor) ~ "Sequence type",
        grepl("^Site", Predictor) ~ "Site",
        grepl("^KL", Predictor) ~ "KL",
        Predictor %in% c("O2","O3","O4", "O5", "Other") ~ "O type",
        TRUE ~ NA_character_
    )) %>% 
    mutate(cat = factor(cat, levels = c("AMR", "KL", "O type", "Sequence type", "Site"))) %>% 
    mutate(signif = case_when(
        P < 0.05 ~ "P < 0.05",
        P >= 0.05 ~ "P ≥ 0.05",
        TRUE ~ NA_character_
    )) %>% 
    arrange(OR) %>% mutate(Predictor = factor(Predictor, levels = unique(Predictor))) %>% 
    ggplot(aes(y = Predictor)) +
    geom_pointrange(aes(x=OR, xmin=lower, xmax=upper, color=signif),
                    position = position_dodge(width=.6), shape=16, size=.85) + 
    scale_color_manual(values = c("darkred", "#FF8A8A"), name=NULL) +
    scale_x_continuous(trans = scales::pseudo_log_trans(base=10),
                       breaks = c(0.1, 1, 2, 5, 10, 50)) +
    geom_vline(xintercept = 1, lty = 2) +
    labs(x = "Adjusted odds ratio with 95% CI (log10 scale)") +
    facet_grid(rows = vars(cat), scales = "free_y", space = "free_y") +
    theme_bw() +
    theme(panel.grid.minor = element_blank(), axis.text = element_text(size=15),
          axis.title = element_text(size=18), strip.text = element_text(size=14),
          legend.text = element_text(size=15), panel.grid.major.y = element_blank(),
          legend.justification = "top", legend.key.height = unit(24, "pt"))
# ggsave(file=file.path(figs_dir, "TransmissionPredictorModel.png"), 
#        plot=trs_model_plot, width=10, height=11) 


############################################
# Use Best or Combined models Table
############################################
# final_trs_model_df %>% 
#     mutate(cat = if_else(Predictor %in% c("ESBL+", "CP+"), "AMR", "Sequence type")) %>%
#     arrange(cat, Predictor) %>% 
#     mutate(OR = sprintf("%.2f", OR),
#            `95% CI` = sprintf("%.2f - %.2f", lower, upper)) %>% 
#     select(-c(lower, upper)) %>% 
#     select(Predictor, Estimate, StdErr, `Adjusted odds ratio` = OR, `95% CI`, P) %>% 
#     write_tsv(., file.path(tables_dir, "TableS7_TransmissionPredictors.tsv"), na="-")


trs_model_AmrSiteStK_df %>% # start with and full-join largest/non-nested models
    mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
    select(Predictor, "AMR + Site + ST + K" = Eff) %>% 
    full_join( # -
        trs_model_AmrSiteStO_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "AMR + Site + ST + O" = Eff),
        by = "Predictor") %>% 
    # left-join smaller nested models
    left_join( # -
        trs_model_Amr_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "AMR" = Eff),
        by = "Predictor") %>% 
    left_join( # -
        trs_model_AmrSite_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "AMR + Site" = Eff),
        by = "Predictor") %>% 
    left_join( # -
        trs_model_AmrSiteSt_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "AMR + Site + ST" = Eff),
        by = "Predictor") %>% 
    left_join( # -
        trs_model_AmrSt_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "AMR + ST" = Eff),
        by = "Predictor") %>% 
    left_join( # -
        trs_model_St_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "ST" = Eff),
        by = "Predictor") %>% 
    left_join( # -
        trs_model_StSite_df %>% 
            mutate(Eff = sprintf("%.1f [%.1f,%.1f], p=%.4f", OR,lower,upper,P)) %>% 
            select(Predictor, "ST + Site" = Eff),
        by = "Predictor") %>%     # view()
    select(Predictor, "AMR", "AMR + Site", "AMR + Site + ST", 
           "AMR + Site + ST + K", "AMR + Site + ST + O") %>% # view() # subset to report 
    write_tsv(., file.path(tables_dir, "TableS7_TransmissionPredictors.tsv"), na="-")


############################################
# # remove redundant models
# rm(list = c(ls(pattern = glob2rx("trs_model_*"))))
```


<br><br>

#### Most common transmission clones (occuring in ≥3 clusters and >1 site)

```{r problem_clones_summary, warning=F, message=F, fig.width=14, fig.height=9}
DT::datatable(top_clones_summary, caption="Clusters summary")

# report top clones
top_clones_summary %>% 
    arrange(n_clusters) %>% 
    mutate(rep=glue("{ST} ({n_clusters} clusters; n={n_isolates_clusters}/{n_isolates})")) %>%
    pull(rep) %>%rev() %>% paste0(collapse = ", ") %>% sub("(.*),", "\\1, and", .)

# Top clones make up what prop of all cases
epi_snp_clusters_wt_upd_STS %>% 
    mutate(ct = if_else(ST %in% top_clones_summary$ST, "Top clone", "Other")) %>% 
    group_by(ct) %>% 
    summarise("n_isolates" = n()) %>% 
    mutate(pct_all_cases = n_isolates/sum(n_isolates) * 100) 
# Top clones make up what prop of cluster cases
epi_snp_clusters_wt_upd_STS %>% 
    filter(!is.na(Cluster)) %>% 
    mutate(ct = if_else(ST %in% top_clones_summary$ST, "Top clone", "Other")) %>% 
    group_by(ct) %>% 
    summarise("n_isolates" = n()) %>% 
    mutate(pct_cluster_cases = n_isolates/sum(n_isolates) * 100) 
# Top clones make up what prop of unclustered cases
epi_snp_clusters_wt_upd_STS %>% 
    filter(is.na(Cluster)) %>% 
    mutate(ct = if_else(ST %in% top_clones_summary$ST, "Top clone", "Other")) %>% 
    group_by(ct) %>% 
    summarise("n_isolates" = n()) %>% 
    mutate(pct_unclustered_cases = n_isolates/sum(n_isolates) * 100) 

# Cluster isolates in top clones
top_clone_isolates_clustered <- epi_snp_clusters_wt_upd_STS %>% 
    filter(!is.na(Cluster)) %>% # only isolates in clusters 
    filter(ST %in% top_clones_summary$ST) %>% # subset top clones
    mutate(ST = factor(ST, levels = top_clones_summary$ST)) %>% 
    select(id, Cluster, ST, Study, Site, K_locus, O_locus, resistance_score,
           Region, Country, Country_Study, Country_Study_Site) %>% 
    mutate(Country_Site = paste0(Country, ": ", Site))
# Cluster isolates in top clones
top_clone_isolates_single <- epi_snp_clusters_wt_upd_STS %>% 
    filter(is.na(Cluster)) %>% # only singleton isolates
    filter(ST %in% top_clones_summary$ST) %>% # subset top clones
    mutate(ST = factor(ST, levels = top_clones_summary$ST)) %>% 
    select(id, Cluster, ST, Study, Site, K_locus, O_locus, resistance_score,
           Region, Country, Country_Study, Country_Study_Site) %>% 
    mutate(Country_Site = paste0(Country, ": ", Site))
# Res score categories of clustered top clone isolates
top_clone_isolates_clustered %>%
    rename_vars_for_plotting(column = "resistance_score") %>%
    count(resistance_score) %>% 
    DT::datatable(., caption="Res score categories of most common transmission clones")
# Res score categories of all clones
epi_snp_clusters %>%
    rename_vars_for_plotting(column = "resistance_score") %>%
    count(resistance_score) 
# How many of the top clone isolates (cluster and singleton) carry at least an ESBL gene
epi_snp_clusters_wt_upd_STS %>% 
    filter(ST %in% top_clones_summary$ST) %>% 
    mutate(g = if_else(is.na(Cluster), "Single", "Cluster")) %>% 
    mutate(r = if_else(resistance_score >= 1, "ESBL", "none")) %>% 
    reframe(n = n(), .by = c(g, r)) %>% 
    mutate(pct = n/sum(n)*100, .by = g) %>% 
    arrange(g)


# order plots
top_clones_order <- top_clones_summary %>% 
    arrange(desc(n_clusters)) %>% pull(ST)
top_clones_order_y <- rev(top_clones_order)
ctry_study_order <- top_clone_isolates_clustered %>% arrange(Region, Country) %>%
    pull(Country_Study) %>% unique()
ctry_site_order <- top_clone_isolates_clustered %>% 
    arrange(Region, Country, as.numeric(gsub("[A-Za-z]| ","",Site))) %>% 
    pull(Country_Site) %>% unique()
ctry_study_site_order <- top_clone_isolates_clustered %>% 
    arrange(Region, Country) %>%
    pull(Country_Study_Site) %>% unique()

top_clones_colors <- setNames(
    c(unname(Polychrome::glasbey.colors(n=15)[-1]), "grey98"),
    c(top_clones_order, "Other STs"))
top_clones_colors['ST35'] <- "#009E60"
top_clones_colors['ST15'] <- "#6F8FBF"

# Transmission profile table of top clones -------------
top_clones_p1_table <- top_clones_summary %>%
    left_join(final_trs_model_df, by=c("ST" = "Predictor")) %>% 
    mutate(ST = factor(ST, levels = top_clones_order_y)) %>% 
    ggplot(aes(y=ST)) +
    geom_text(aes(x=0, label=n_clusters), hjust=0, size=4) +
    annotate("text", x=0, y=nrow(top_clones_summary)+1, label="Clusters", 
             hjust=0, fontface="bold", size=4) +
    geom_vline(xintercept = 0.35, color = "grey80") +
    geom_text(aes(x=0.37, label=n_isolates_singletons), hjust=0, size=4) +
    annotate("text", x=0.37, y=nrow(top_clones_summary)+1, label="Single", 
             hjust=0, fontface="bold", size=4) +
    geom_vline(xintercept = 0.66, color = "grey80") +
    geom_text(aes(x=0.68, label=n_sites), hjust=0, size=4) +
    annotate("text", x=0.68, y=nrow(top_clones_summary)+1, label="Sites",
             hjust=0, fontface="bold", size=4) +
    geom_vline(xintercept = 0.91, color = "grey80") +
    geom_text(aes(x=0.93, label=n_sites_clusters), hjust=0, size=4) +
    annotate("text", x=0.93, y=nrow(top_clones_summary)+1, label="Sites\n(clusters)", 
             hjust=0, vjust=0.2, fontface="bold", size=4) +
    geom_vline(xintercept = 1.30, color = "grey80") +
    geom_text(aes(x=1.32, label =sprintf('%.1f [%.1f-%.1f]', OR,lower,upper)), 
              hjust=0, size=4) +
    annotate("text", x=1.32, y=nrow(top_clones_summary)+1, label="aOR [95% CI]",
             hjust=0, fontface="bold", size=4) +
    geom_vline(xintercept = 1.83, color = "grey80") +
    geom_text(aes(x=1.85, fontface = ifelse(P < 0.05, "bold", "plain"),
                  label = ifelse(P > 0.099, sprintf('%.2f', P), sprintf('%.3f', P))), 
              hjust=0, size=4) +
    annotate("text", x=1.85, y=nrow(top_clones_summary)+1, label="P", 
             hjust=0, fontface="bold.italic", size=4) +
    labs(title = "Transmission profile") +
    geom_hline(yintercept = nrow(top_clones_summary)+0.5) +
    geom_vline(xintercept = -0.05) +
    scale_x_continuous(expand = expansion(mult = c(0.005, 0.5))) +
    theme_void() + custom_plots_theme +
    theme(axis.text.x = element_blank(), axis.title = element_blank(),
          axis.ticks.x = element_blank(), panel.grid = element_blank()) +
    coord_cartesian(ylim = c(1, nrow(top_clones_summary)+1), xlim=c(-0.1,1.4), clip="off")


# top clones by n_clusters ------------------------
top_clones_p1 <- top_clones_summary %>%
    mutate(ST = factor(ST, levels = top_clones_order_y)) %>%
    ggplot2::ggplot(aes(x=ST, y = n_clusters)) +
    ggplot2::geom_col(fill = "#0571B0") +
    ggplot2::labs(y = "Number of clusters", fill = "Median cluster size",
                  title = "# clusters") +
    ggplot2::theme_bw() +
    ggplot2::scale_y_continuous(breaks = seq(0, max(top_clones_summary$n_clusters), 4)) +
    custom_plots_theme + 
    ggplot2::theme(axis.title = element_blank(),
                   panel.grid.major.x = element_blank(), 
                   panel.grid.minor.x = element_blank()) +
    ggplot2::coord_flip()

# n_sites_topclones -------------------------
top_clones_p2 <- top_clones_summary %>%
    mutate(ST = factor(ST, levels = top_clones_order_y)) %>% 
    ggplot2::ggplot(aes(x=ST, y = n_sites_clusters)) +
    ggplot2::geom_col(fill = "#0571B0") +
    ggplot2::labs(y = "Number of sites", title = "# sites") +
    ggplot2::theme_bw() +
    ggplot2::scale_y_continuous(breaks = seq(0, max(top_clones_summary$n_isolates_clusters), 2)) +
    custom_plots_theme + 
    theme(axis.title = element_blank(),
          panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
    ggplot2::coord_flip() + 
    ggplot2::guides(fill = "none")

## site_distribution_top_clones; coded sites colored by region -------------
top_clones_p3 <- top_clone_isolates_clustered %>% group_by(ST) %>% 
    add_count(ST, Site, name = "n_isolates") %>% 
    mutate(ST = factor(ST, levels=top_clones_order_y)) %>% 
    mutate(Country_Site = factor(Country_Site, levels=ctry_site_order)) %>%
    mutate(text = paste("Site: ", Country_Site, 
                        "\nST: ", ST, 
                        "\nN cluster isolates: ", n_isolates)) %>% 
    ggplot2::ggplot(aes(y = ST, x = Country_Site, size = n_isolates, 
                        fill = Region, text = text)) +
    ggplot2::geom_point(shape=21, color='black', stroke=0.1) +
    scale_size_continuous(range=c(2, 10), name = "No. cluster isolates",
                          guide = guide_legend(override.aes = list(stroke = 0.8), order=1)) +
    ggplot2::scale_fill_manual(values = region_colors,
                                guide=guide_legend(override.aes=list(size=6), , order=2)) +
    ggplot2::theme_classic() +
    ggplot2::labs(title = "Site distribution", y = "Sites") +
    custom_plots_theme + 
    ggplot2::theme(axis.text.x = element_text(size=8, angle=90, hjust=1, vjust=0.5),
                   axis.title = element_blank(), panel.grid.major.x = element_line(color="grey95"),
                   panel.grid.major.y = element_line(colour="grey70"), 
                   panel.grid.minor.x = element_blank()) 

# AMR_distribution_top_clones ----------------------
top_clones_p4 <- top_clone_isolates_clustered %>% 
    rename_vars_for_plotting(column = "resistance_score") %>% 
    mutate(ST = factor(ST, levels = top_clones_order_y)) %>% 
    mutate(resistance_score = factor(resistance_score, levels = rev(c(
        '0: ESBL-, Carb-', '1: ESBL+, Carb-', '2: Carb+', '3: Carb+, Col+'
    )))) %>% 
    ggplot(aes(y = ST, fill = resistance_score)) +
    geom_bar(position="fill") +
    scale_fill_manual(values = res_colors) +
    guides(fill = guide_legend(reverse = TRUE)) +
    theme_classic() +
    custom_plots_theme +
    labs(title = "Resistance distribution", fill = "Resistance score") +
    theme(axis.title = element_blank(),
          panel.grid.major.y = element_line(colour="grey70"),
          panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())  

# STs making up large clusters
STs_by_cluster_size <- all_clusters_info_wt_upd_STs %>% 
    mutate(Cluster_size = `N isolates`) %>% 
    arrange(desc(Cluster_size)) %>% 
    mutate(Cluster_size = factor(Cluster_size, levels = unique(Cluster_size))) %>% 
    mutate(top_ST = if_else(ST %in% top_clones_order, ST, "Other STs")) %>% 
    mutate(top_ST = factor(top_ST, levels = c(top_clones_order, "Other STs"))) %>% 
    select(Cluster_size, `N isolates`, top_ST, ST, everything())

# Plot (bars)
STs_by_cluster_size_plot <- STs_by_cluster_size %>% 
    mutate(Cluster_size = factor(Cluster_size, levels = rev(unique(Cluster_size)))) %>%
    ggplot(aes(x = Cluster_size, y = `N isolates`, fill=top_ST)) +
    scale_fill_manual(values=top_clones_colors, name="ST") +
    geom_col(color="black") +
    labs(x = "Cluster size", y = "No. isolates",
         title = "STs associated with large clusters") +
    theme_classic() + custom_plots_theme +
    theme(panel.grid.major.y = element_line(colour="grey90"),
          panel.grid.major.x = element_line(colour="grey90"),
          axis.text.x = element_text(angle=45, hjust=1)) +
    guides(fill = guide_legend(ncol=1))
# Plot (bubbles)
STs_by_cluster_size_bubbles_plot <- STs_by_cluster_size %>% 
    mutate(top_ST = factor(top_ST, levels = rev(levels(top_ST)))) %>% 
    mutate(Cluster_size = factor(Cluster_size, levels = rev(unique(Cluster_size)))) %>%
    ggplot(aes(x = Cluster_size, y = top_ST, fill=top_ST, size = `N isolates`)) +
    geom_hline(yintercept = seq(0.5, length(unique(STs_by_cluster_size$top_ST))+0.5, 1), 
               linewidth=0.25, color="grey90") +
    geom_vline(xintercept =seq(1.5, length(unique(STs_by_cluster_size$Cluster_size)), 1), 
               linewidth=0.25, color="grey90") +
    geom_point(color="black", shape=21, stroke=0.2) +
    scale_fill_manual(values=top_clones_colors, name="ST") +
    scale_size_continuous(range=c(2,12), breaks = c(10, 50, 100, 150),
                          guide = guide_legend(override.aes = list(stroke = 0.8))) +
    labs(x = "Cluster size", y = NULL, size = "No. isolates",
         title = "STs associated with large clusters") +
    guides(fill = "none") +
    theme_classic() + custom_plots_theme +
    theme(axis.text.x = element_text(angle=45, hjust=1),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.ticks.y = element_blank(),
          legend.key.width = unit(5, "pt")) +
    coord_cartesian(clip = "off")

# Non-top STs making up large clusters
STs_by_cluster_size %>% filter(top_ST == "Other STs")

# What prop of large clusters (N >= ) are made up by most commonly transmitted STs
STs_by_cluster_size %>% 
    mutate(ST_grp = if_else(top_ST == "Other STs", "Others", "Top ST")) %>% 
    filter(`N isolates` >= 15) %>% count(ST_grp) %>% 
    mutate(pct=round(n/sum(n)*100,1)) 

# Combine top clones plots ----------------------
top_clones_combined_plot <-
    (top_clones_p1_table  + theme(plot.title = element_text(vjust = 3)) |
        top_clones_p3 + theme(axis.text.y = element_blank(), axis.text.x = element_text(size=10)) +
            coord_cartesian(ylim = c(1, nrow(top_clones_summary)+1)) |
        top_clones_p4 + theme(axis.text.y = element_blank()) +
            coord_cartesian(ylim = c(1, nrow(top_clones_summary)+1))
     ) +
    plot_annotation(
        title = 'Commonly transmitted STs', subtitle = 'STs in ≥3 clusters & ≥2 sites',
        theme = theme(plot.title = element_text(size = 16), plot.title.position = "panel"),
        tag_levels = "A", tag_prefix = "(", tag_suffix = ")") +
    plot_layout(widths=c(1.8,1.8,0.8), guides="collect") &
    theme(plot.tag = element_text(face="bold", vjust=-3.3))

top_clones_combined_plot2 <- 
    ( top_clones_combined_plot + STs_by_cluster_size_bubbles_plot ) + 
    plot_layout(
        design = c(area(1,1,1,4), area(1,5,1,7), area(1,8,1,9), area(2,1,2,7)),  
        nrow = 2, guides = "collect") +
    plot_annotation(theme = theme(plot.title = element_text(size = 20),
                                  plot.subtitle = element_text(size = 14))) & 
    theme(plot.tag = element_text(face="bold", vjust=-3.3),
          legend.justification = "top")
ggsave(file=file.path(figs_dir, "Figure2_TransmissionClones.svg"),
       plot=top_clones_combined_plot2, width=15, height=12)

```


```{r introductions_vs_clusters}
########################################################################################
# Compare No. introductions to No. clusters for each ST
########################################################################################
intros_vs_clusters <- epi_snp_clusters_dereplicated %>% 
    group_by(ST) %>% 
    reframe(intros = n(), clusters = n_distinct(Cluster, na.rm=T)) %>% 
    mutate(ST_grp = if_else(ST %in% top_clones_summary$ST, "Common ST", "Other STs")) %>% 
    mutate(common_ST = if_else(ST %in% top_clones_summary$ST, ST, "Other STs")) %>% 
    mutate(common_ST = factor(common_ST, levels = c(top_clones_summary$ST, "Other STs"))) %>% 
    left_join(final_trs_model_df %>% select(Predictor, P), by = c("ST" = "Predictor")) %>% 
    mutate(Pgrp = case_when(
        P < 0.05 ~ "p<0.05",
        P >= 0.05 ~ "p≥0.05",
        TRUE ~ NA_character_
    )) %>% 
    mutate(signif = if_else(!is.na(Pgrp), paste0(ST_grp, ", ", Pgrp), ST_grp)) %>% 
    arrange(P) %>% mutate(signif = factor(signif, levels = unique(signif))) %>% 
    arrange(desc(intros)) %>% 
    mutate(pct_clustered = round(clusters/intros*100,1)) %>% 
    arrange(desc(pct_clustered))

# crude transmission probability 
# all
intros_vs_clusters %>% 
    reframe(trs_prob = paste0(mean(pct_clustered, na.rm = T), " [",
                              paste0(range(pct_clustered, na.rm=T), collapse = " - "), "]"))
# in each group
intros_vs_clusters %>% # filter(common_ST != "Other STs") %>% 
    group_by(ST_grp) %>%
    reframe(trs_prob = paste0(mean(pct_clustered, na.rm = T), " [", 
                              paste0(range(pct_clustered, na.rm=T), collapse = " - "), "]"))

# Are the fractions significantly different between common STs and others
intros_vs_clusters %>%
    group_by(ST_grp) %>%
    reframe(cl = sum(clusters), sing = sum(intros - clusters)) %>% 
    column_to_rownames("ST_grp") %>% as.matrix() %>%
    prop.test()



# Plot intros and clusters
intros_vs_clusters_plot <- intros_vs_clusters %>% 
    left_join(clones_summary %>% select(ST, n_sites_clusters), by = "ST") %>% 
    arrange(desc(P)) %>% 
    ggplot(aes(x=intros, y=clusters)) +
    geom_abline(color="grey30", linetype = 2, linewidth = 1) +
    geom_point(aes(color = signif, size = n_sites_clusters), alpha=.8) +
    scale_color_manual(values = c("#750000", "#FF8A8A", "grey70"), name="ST") +
    scale_size_continuous(
            name="No. of sites (cluster cases)", range=c(0.5,6),
            breaks=function(x) unique(floor(pretty(seq(min(x), (max(x) + 1) * 1.1))))) +
    geom_smooth(method=lm, se=FALSE, formula=y~x-1, color="grey30") +
    ggtext::geom_richtext(
        x=32, y=15, size=3.5, label.colour=NA, fill=NA, hjust=0,
        data = data.frame(intros=1, clusters=1), # 1 row to plot once
        label = get_lm_eqn(intros_vs_clusters %>% select(x=intros, y=clusters), format = T)) +
    ggrepel::geom_text_repel(
        data = intros_vs_clusters %>% filter(common_ST != "Other STs"),
        aes(label=common_ST, color = signif), max.overlaps=40, size=3.5, seed=5, show.legend=F) +
    labs(x = "No. of unique introductions", y = "No. of clusters") +
    theme_bw() + custom_plots_theme +
    guides(color = guide_legend(override.aes = list(size=7), order=2,
                                theme = theme(legend.key.spacing.x = unit(6, "mm"))),
           size = guide_legend(nrow=1, order=1, theme = theme(legend.key.spacing.x = unit(4, "mm")))) +
    # guides(color = "none") +
    theme(axis.text=element_text(size=12), axis.title = element_text(size=14),
          plot.title = element_text(size=14), legend.position = "bottom",
          legend.direction = "horizontal", legend.box = "vertical",
          legend.margin = margin(b=0.1, unit="mm"),
          legend.text = element_text(size=12, margin = margin(l=0.1, r=2, unit = "mm")), 
          legend.title = element_text(size=12, margin = margin(l=0, r=4, unit = "mm"))) 



# Plot distribution of fractions of clusters/introductions
intros_clusters_frac_distrib_plot <- intros_vs_clusters %>% 
    ggplot(aes(x = pct_clustered/100)) +
    geom_histogram(aes(fill = common_ST), binwidth = 0.05, color="black", size=0.25) +
    scale_fill_manual(values = top_clones_colors, name=NULL) +
    labs(x = "Fraction of introductions resulting in clusters", y = "No. of STs") +
    theme_bw() + custom_plots_theme +
    theme(axis.text=element_text(size=12), axis.title = element_text(size=14),
          plot.title = element_text(size=14),
          legend.text = element_text(size=12, margin = margin(l=0.5, r=0.5, unit = "mm")), 
          legend.title = element_text(size=14, margin = margin(t=4, b=1, unit = "mm"))) +
    ggbreak::scale_y_break(c(15, 105), ticklabels = c(110, 115))


# Plot fraction of ESBL/CP cases for singletons vs clusters
ESBL_CP_frac_plot <- epi_snp_clusters_dereplicated %>% 
    # filter(ST %in% top_clones_order) %>%
    mutate(case_type = if_else(is.na(Cluster), "Single", "Cluster")) %>% 
    mutate(hasESBLorCP = if_else(resistance_score >= 1, "Yes", "No")) %>% 
    group_by(ST, case_type) %>% count(hasESBLorCP) %>% 
    group_by(ST, case_type) %>% 
    mutate(total = sum(n), fracWithESBLorCP = n/total) %>% 
    ungroup() %>% 
    pivot_wider(id_cols = c(ST, case_type), names_from = hasESBLorCP, 
                values_from = fracWithESBLorCP, values_fill = 0) %>% 
    pivot_wider(id_cols = ST, names_from = case_type, values_from = Yes) %>% 
    filter(!is.na(Cluster)) %>% 
    left_join(intros_vs_clusters %>% select(ST, signif), by="ST") %>% 
    arrange(desc(signif)) %>% # use levels control order of points
    # mutate(ST = factor(ST, levels = ST)) %>% 
    ggplot(aes(x = Single, y = Cluster)) +
    geom_point(aes(fill = signif), size=6, color="black", stroke=0.25, 
               shape=21, alpha=0.7, position=position_jitter(width=0.005, height=0.005)) +
    scale_fill_manual(values = c("#750000", "#FF8A8A", "grey70"), name="ST",
                       guide = guide_legend(override.aes = list(alpha = 1))) +
    labs(x = "ESBL/CP fraction among singletons", y = "ESBL/CP fraction among clusters") +
    theme_bw() + custom_plots_theme +
    guides(fill = guide_legend(override.aes = list(size=7), order=2,
                                theme = theme(legend.key.spacing.x = unit(6, "mm"))),
           size = guide_legend(nrow=1, order=1, theme = theme(legend.key.spacing.x = unit(4, "mm")))) +
    guides(fill = "none") + # dont show, duplicated in combined plot
    theme(axis.text=element_text(size=12), axis.title = element_text(size=14),
          plot.title = element_text(size=14), legend.position = "bottom",
          legend.direction = "horizontal", legend.box = "vertical",
          legend.text = element_text(size=12, margin = margin(l=0.1, r=2, unit = "mm")), 
          legend.title = element_text(size=12, margin = margin(l=0, r=4, unit = "mm"))) +
    coord_cartesian(clip = "off", xlim=c(0,1), ylim=c(0,1))
    

# Combine plots
intros_clusters_combined_plot <- (
    intros_clusters_frac_distrib_plot + 
    intros_vs_clusters_plot + 
    ESBL_CP_frac_plot
    ) + 
    plot_layout(design = c(area(1,1,4,7), area(5,1,9,4), area(5,5,8,8))) +
    plot_annotation(tag_levels = "A", tag_prefix = "(", tag_suffix = ")",
                    theme = theme(plot.title = element_text(size = 20))) & 
    theme(plot.tag = element_text(face="bold"), legend.justification = "top")

ggsave(file=file.path(figs_dir, "Figure3_IntrosVsClusters.png"),
       plot=intros_clusters_combined_plot, width=14, height=10)
ggsave(file=file.path(figs_dir, "Figure3_IntrosVsClusters.tif"),
       plot=intros_clusters_combined_plot, width=14, height=10)


########################################################################################
# Binomial tests of association between common STs and transmission
########################################################################################
intros_vs_clusters_top <- intros_vs_clusters %>% filter(ST_grp == "Common ST")
intros_vs_clusters_top$binom.p <- sapply(1:nrow(intros_vs_clusters_top), function(i) {
  binom.test(x = intros_vs_clusters_top$clusters[i], intros_vs_clusters_top$intros[i], 
             alternative = "greater")$p.value
})
intros_vs_clusters_top$binom.padj <- p.adjust(intros_vs_clusters_top$binom.p, method = "bonferroni")
intros_vs_clusters_top$binom.signif <- intros_vs_clusters_top$binom.padj < 0.05
```


<br><br>

### Resistance distribution  

```{r resistance_distro, fig.width=10, fig.height=7}
# Resistance score categories all clones
epi_snp_clusters_wt_upd_STS %>%
    rename_vars_for_plotting(column = "resistance_score") %>%
    count(resistance_score) %>%
    DT::datatable(., caption="Res score categories of all clones")

res_score_df <- epi_snp_clusters_wt_upd_STS %>% 
    mutate(case_type = if_else(is.na(Cluster), "Single", "Cluster")) %>% 
    add_count(Site, name="n_isolates_per_site") %>% 
    mutate(m_Site = paste0(Study_lab, "\n", coded_site, " (N=", n_isolates_per_site,")")) %>% 
    mutate(resistance_score_ori = as.numeric(resistance_score)) %>% 
    rename_vars_for_plotting(column = "resistance_score") %>% 
    arrange(Study, coded_site) %>% 
    mutate(m_Site = factor(m_Site, levels = unique(m_Site))) %>%
    arrange(Study, desc(resistance_score)) %>% 
    mutate(clone_type = if_else(ST %in% top_clones_summary$ST, "Most common", "Others")) %>%
    select(id, Cluster, Study, Study_lab, Country, m_Site, Country_Study_Site, coded_site, 
           ST, resistance_score, resistance_score_ori, case_type, clone_type)

# De-replicate clusters before comparison of res scores
# i.e., 1 sample per cluster with consensus res score
res_score_df_dereplicated <- res_score_df %>% 
    # dereplicate clusters
    filter(!is.na(Cluster)) %>% 
    group_by(Cluster) %>% 
    add_count(resistance_score) %>% 
    slice_max(order_by = n, with_ties = F) %>% ungroup() %>% 
    # add in non-cluster isolates
    bind_rows(res_score_df %>% filter(is.na(Cluster))) %>% 
    add_count(coded_site, name="n_isolates_per_site") %>% 
    mutate(m_Site = paste0(Study_lab, "\n", coded_site, " (N=", n_isolates_per_site,")")) %>%
    arrange(Study, coded_site) %>%
    mutate(m_Site = factor(m_Site, levels = unique(m_Site)))

```


```{r ESBL_and_CP_by_case_type}
res_case_type_plot <- res_score_df_dereplicated %>% 
    rbind(res_score_df_dereplicated %>% 
              mutate(m_Site=paste0("ALL SITES\n(N=", nrow(res_score_df_dereplicated), ")"))) %>% 
    mutate(resistance_score = factor(resistance_score, levels = rev(names(res_colors)))) %>% 
    ggplot2::ggplot(aes(x = case_type, fill = resistance_score)) +
    ggplot2::geom_bar(position="fill") +
    geom_text(aes(label=after_stat(count), color=resistance_score), stat="count", 
              position=position_fill(vjust=0.5), size=3.5) +
    ggplot2::scale_fill_manual(values = res_colors) +
    ggplot2::scale_color_manual(values=c("orange", "black","black","black")) +
    facet_wrap(~m_Site, ncol=6) + 
    ggplot2::theme_bw() +
    custom_plots_theme +
    ggplot2::labs(fill="Resistance score", x="Case type", y="Proportion") +
    ggplot2::theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(),
                   legend.position = "bottom", axis.title = element_text(size=12),
                   axis.text=element_text(size=10), legend.key.spacing.x = unit(10, "mm"),
                   legend.text=element_text(size=10, margin=margin(l=0.5, r=0.5, unit="mm")),
                   legend.title=element_text(size=10), strip.text=element_text(size=10)) +
    ggplot2::guides(color = "none", fill = guide_legend(reverse = TRUE)) 
# ggsave(file=file.path(figs_dir, "FigureSX_Resistance_byCaseType.png"), 
#        plot=res_case_type_plot, width=10, height=11)
# ggsave(file=file.path(figs_dir, "FigureSX_Resistance_byCaseType.tif"), 
#        plot=res_case_type_plot, width=10, height=11)
# ggsave(file=file.path(figs_dir, "FigureSX_Resistance_byCaseType.svg"), 
#        plot=res_case_type_plot, width=10, height=11)
res_case_type_plot


# Compare ESBL proportions by case type
ESBL_by_case_type_contige_table <- res_score_df_dereplicated %>%
    mutate(ESBL = if_else(resistance_score_ori >= 1, "ESBL", "non-ESBL")) %>%
    select(case_type, ESBL) %>% table()
ESBL_by_case_type_test <- prop.test(ESBL_by_case_type_contige_table)
ESBL_by_case_type <- res_score_df_dereplicated %>% 
    mutate(Category = if_else(resistance_score_ori >= 1, "Yes", "No")) %>% 
    group_by(case_type, Category) %>% 
    summarise(N = n()) %>% 
    mutate(total_isolates = sum(N, na.rm = TRUE),  
           pct = round((N / total_isolates) * 100, 1)) %>% 
    select(-total_isolates) %>% ungroup() %>%  
    mutate(Val = paste0(N, " (", pct, "%)")) %>% 
    pivot_wider(id_cols = Category, names_from = case_type, values_from = Val) %>% 
    mutate(Test="ESBL presence", pval=ESBL_by_case_type_test$p.value) %>% 
    relocate(Test, .before = "Category")
ESBL_by_case_type

# Compare Carba proportions by case type
CP_by_case_type_contige_table <- res_score_df_dereplicated %>%
    mutate(CP = if_else(resistance_score_ori >= 2, "Carba", "non-carba")) %>%
    select(case_type, CP) %>% table()
CP_by_case_type_test <- prop.test(CP_by_case_type_contige_table)
CP_by_case_type <- res_score_df_dereplicated %>% 
    mutate(Category = if_else(resistance_score_ori >= 2, "Yes", "No")) %>% 
    group_by(case_type, Category) %>% 
    summarise(N = n()) %>% 
    mutate(total_isolates = sum(N, na.rm = TRUE),  
           pct = round((N / total_isolates) * 100, 1)) %>% 
    select(-total_isolates) %>% ungroup() %>%  
    mutate(Val = paste0(N, " (", pct, "%)")) %>% 
    pivot_wider(id_cols = Category, names_from = case_type, values_from = Val) %>% 
    mutate(Test="Carbapenem presence", pval=CP_by_case_type_test$p.value) %>% 
    relocate(Test, .before = "Category")
CP_by_case_type

# Combine dfs
bind_rows(ESBL_by_case_type, CP_by_case_type) %>% 
    write_tsv(., file.path(tables_dir, "TableS6_ResByCaseType.tsv"))
```


```{r ESBL_and_CP_by_clone_type, fig.width=8, fig.height=6}
# Compare resistance scores. Top clones vs Others
res_score_df_dereplicated %>% group_by(clone_type) %>% reframe(n_distinct(ST))

# ESBL proportions by clone type
ESBL_by_clone_type_contige_table <- res_score_df_dereplicated %>%
    mutate(ESBL = if_else(resistance_score_ori >= 1, "ESBL", "non-ESBL")) %>%
    select(clone_type, ESBL) %>% table()
ESBL_by_clone_type_test <- prop.test(ESBL_by_clone_type_contige_table)
ESBL_by_clone_type <- res_score_df_dereplicated %>% 
    mutate(Category = if_else(resistance_score_ori >= 1, "Yes", "No")) %>% 
    group_by(clone_type, Category) %>% 
    summarise(N = n()) %>% 
    mutate(total_isolates = sum(N, na.rm = TRUE),  
           pct = round((N / total_isolates) * 100, 1)) %>% 
    select(-total_isolates) %>% ungroup() %>%  
    mutate(Val = paste0(N, " (", pct, "%)")) %>% 
    pivot_wider(id_cols = Category, names_from = clone_type, values_from = Val) %>% 
    mutate(Test="ESBL presence", pval=ESBL_by_clone_type_test$p.value) %>% 
    relocate(Test, .before = "Category")
ESBL_by_clone_type


# Carba proportions by clone type
CP_by_clone_type_contige_table <- res_score_df_dereplicated %>%
    mutate(CP = if_else(resistance_score_ori >= 2, "Carba", "non-carba")) %>%
    select(clone_type, CP) %>% table()
CP_by_clone_type_test <- prop.test(CP_by_clone_type_contige_table)
CP_by_clone_type <- res_score_df_dereplicated %>% 
    mutate(Category = if_else(resistance_score_ori >= 2, "Yes", "No")) %>% 
    group_by(clone_type, Category) %>% 
    summarise(N = n()) %>% 
    mutate(total_isolates = sum(N, na.rm = TRUE),  
           pct = round((N / total_isolates) * 100, 1)) %>% 
    select(-total_isolates) %>% ungroup() %>%  
    mutate(Val = paste0(N, " (", pct, "%)")) %>% 
    pivot_wider(id_cols = Category, names_from = clone_type, values_from = Val) %>% 
    mutate(Test="Carbapenem presence", pval=CP_by_clone_type_test$p.value) %>% 
    relocate(Test, .before = "Category")
CP_by_clone_type

bind_rows(ESBL_by_clone_type, CP_by_clone_type) %>% 
    write_tsv(., file.path(tables_dir, "TableS8_ResByCloneType.tsv"))


```


<br><br>

## Effect of distance method on cluster estimates

<details>
<summary>Click to expand</summary> 
```{r SKA_data}
# BARNARDS
BARNARDS_ska_snp_and_epi_data <- read_csv("data/BARNARDS/BARNARDS_pairwise_snpepi_data_SKA.csv", show_col_types=F) 
BARNARDS_epi_snp_clusters_ska <- get_clusters(
        BARNARDS_ska_snp_and_epi_data, snp_distance_threshold, 
        temporal_distance_threshold, BARNARDS_data
    ) %>%
    mutate(Cluster=if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster))) %>% 
    mutate(dist_method = "SKA")

# SPINZ
SPINZ_ska_snp_and_epi_data <- read_csv("data/SPINZ/SPINZ_pairwise_snpepi_data_SKA.csv", show_col_types=F) 
SPINZ_epi_snp_clusters_ska <- get_clusters(
        SPINZ_ska_snp_and_epi_data, genet_dist=snp_distance_threshold,
        temp_dist=temporal_distance_threshold, metadata=SPINZ_data
    ) %>% 
    mutate(Cluster = if_else(is.na(Cluster), Cluster, paste0(Study, '_', Cluster))) %>% 
    mutate(dist_method = "SKA")

```
</details>


<br>


#### Compare estimates at variable SNP and temporal distance
```{r dist_method_comp_variable_snp, fig.height=7, fig.width=12}
# BARNARDS SKA
BARNARDS_ska_sens_df <- get_cluster_sensitivity(BARNARDS_ska_snp_and_epi_data, BARNARDS_data,
                                                sens_snp_range, sens_temp_range)

# SPINZ SKA
SPINZ_ska_sens_df <- get_cluster_sensitivity(SPINZ_ska_snp_and_epi_data, SPINZ_data,
                                            sens_snp_range, sens_temp_range)

# SPINZ and BARNARDS -  SKA and PW sensisitivity
combined_ska_pw_sens_df <- bind_rows(
    BARNARDS_sens_df %>% mutate(Group = "BARNARDS (PW)"),
    BARNARDS_ska_sens_df %>% mutate(Group = "BARNARDS (SKA)"),
    SPINZ_sens_df %>% mutate(Group = "SPINZ (PW)"),
    SPINZ_ska_sens_df %>% mutate(Group = "SPINZ (SKA)")
)

# Sensitivity plots
dist_method_sens_clust_plot <- plot_sensitivity_SNP_vs_temp_range_2(
    combined_ska_pw_sens_df, temp_dist_vals = sens_temp_dist_vals,
    prop_var = "cluster_prop", y_title = "Proportion in clusters",
    plot_title = "(A) Cluster proportion", selected_dist_thresh=snp_distance_threshold,
    selected_temp_thresh=temporal_distance_threshold, facet_col = "Group") +
    theme(axis.text = element_text(size=12), axis.title = element_text(size=14),
          strip.text = element_text(size = 12), plot.title = element_text(face="bold"))

dist_method_sens_trans_plot <- plot_sensitivity_SNP_vs_temp_range_2(
    combined_ska_pw_sens_df, temp_dist_vals = sens_temp_dist_vals,
    prop_var = "transmission_prop", y_title = "Proportion due to\ntransmission",
    plot_title = "(B) Transmission proportion", selected_dist_thresh=snp_distance_threshold,
    range_cols=c("white", "navyblue", "lightblue"),
    selected_temp_thresh=temporal_distance_threshold, facet_col = "Group") +
    theme(axis.text = element_text(size=12), axis.title = element_text(size=14),
          strip.text = element_text(size = 12), plot.title = element_text(face="bold"))

# Pairwise distance comparison plots
# BARNARDS
SKA_PW_dists_BARNARDS <- BARNARDS_snp_and_epi_data %>% 
    mutate(L1 = if_else(iso1 < iso2, iso1, iso2)) %>% 
    mutate(L2 = if_else(iso1 > iso2, iso1, iso2)) %>% 
    select(L1, L2, Pathogenwatch = dist) %>% 
    left_join(BARNARDS_ska_snp_and_epi_data %>% select(L1, L2, SKA = dist),
              by = c("L1", "L2")) 
SKA_PW_dists_BARNARDS_plot_100 <- SKA_PW_dists_BARNARDS %>% 
    filter(Pathogenwatch <= 100 & SKA <= 100) %>% 
    group_by(Pathogenwatch, SKA) %>% reframe(Pairs = n()) %>% 
    ggplot(aes(y = SKA, x = Pathogenwatch, size = Pairs)) + 
    geom_abline(color="grey80") +
    geom_point() +
    labs(title = NULL, x=NULL, y=NULL) +
    theme_bw() +
    coord_equal(xlim=c(0,100), ylim=c(0,100)) +
    theme(axis.text = element_text(size=10), plot.margin = unit(c(0, 0, 0, 0), "mm"),
          legend.title = element_text(size=10), legend.text = element_text(size=10))
SKA_PW_dists_BARNARDS_plot <- SKA_PW_dists_BARNARDS %>% 
    ggplot(aes(y = SKA, x = Pathogenwatch)) + 
    labs(title = "(C) Pairwise genetic distances (BARNARDS)") +
    geom_point() + theme_bw() +
    theme(axis.text = element_text(size=12), axis.title = element_text(size=14), 
          plot.title = element_text(face="bold")) +
    patchwork::inset_element(SKA_PW_dists_BARNARDS_plot_100 + theme(legend.position = "none") +
                                 scale_size_continuous(range=c(0.05,2)),
                             0.6, 0.01, 0.99, 0.65, ignore_tag=T)

# SPINZ
SKA_PW_dists_SPINZ <- SPINZ_snp_and_epi_data %>% 
    mutate(L1 = if_else(iso1 < iso2, iso1, iso2)) %>% 
    mutate(L2 = if_else(iso1 > iso2, iso1, iso2)) %>% 
    select(L1, L2, Pathogenwatch = dist) %>% 
    left_join(SPINZ_ska_snp_and_epi_data %>% select(L1, L2, SKA = dist),
              by = c("L1", "L2")) 
SKA_PW_dists_SPINZ_plot_100 <- SKA_PW_dists_SPINZ %>% 
    filter(Pathogenwatch <= 100 & SKA <= 100) %>% 
    group_by(Pathogenwatch, SKA) %>% reframe(Pairs = n()) %>% 
    ggplot(aes(y = SKA, x = Pathogenwatch, size = Pairs)) + 
    geom_abline(color="grey80") +
    geom_point() +
    labs(title = NULL, x=NULL, y=NULL) +
    theme_bw() +
    coord_equal(xlim=c(0,100), ylim=c(0,100)) +
    theme(axis.text = element_text(size=10), plot.margin = unit(c(0, 0, 0, 0), "mm"),
          legend.title = element_text(size=10), legend.text = element_text(size=10))
SKA_PW_dists_SPINZ_plot <- SKA_PW_dists_SPINZ %>% 
    ggplot(aes(y = SKA, x = Pathogenwatch)) + 
    labs(title = "(D) Pairwise genetic distances (SPINZ)") +
    geom_point() + theme_bw() +
    theme(axis.text = element_text(size=12), axis.title = element_text(size=14), 
          plot.title = element_text(face="bold")) + 
    patchwork::inset_element(SKA_PW_dists_SPINZ_plot_100 + theme(legend.position = "none") +
                                 scale_size_continuous(range=c(0.05,2)),
                             0.6, 0.01, 0.99, 0.65, ignore_tag=T)


# Combine plots
dist_method_sens_plots <- dist_method_sens_clust_plot + sens_clust_legend + 
    dist_method_sens_trans_plot + sens_trans_legend + 
    SKA_PW_dists_BARNARDS_plot +
    SKA_PW_dists_SPINZ_plot + 
    patchwork::plot_layout(
        design = c(patchwork::area(1,1,1,3), patchwork::area(1,4),
                   patchwork::area(2,1,2,3), patchwork::area(2,4),
                   patchwork::area(3,1), patchwork::area(3,3)), 
        guides = "collect", widths = unit(c(1,0.3,1,0.3,1,1), "null"))
dist_method_sens_plots
ggsave(file=file.path(figs_dir, "AppendixS1Fig4_DistanceMethodSensitivity.png"), 
       plot=dist_method_sens_plots, width=12, height=10, bg="white")
ggsave(file=file.path(figs_dir, "AppendixS1Fig4_DistanceMethodSensitivity.tif"), 
       plot=dist_method_sens_plots, width=12, height=10, bg="white")

# Summary
combined_ska_pw_sens_df %>% 
    filter(distance_threshold %in% c(10), temporal_threshold %in% c(2,4,8)) %>%
    mutate(temporal_threshold = paste0("temp_",temporal_threshold)) %>% 
    pivot_wider(id_cols=Group, names_from=temporal_threshold, values_from=cluster_prop) %>% 
    mutate(cluster_prop = sprintf("%.2f [%.2f-%.2f]", temp_4, temp_2, temp_8))
combined_ska_pw_sens_df %>% 
    filter(distance_threshold %in% c(10), temporal_threshold %in% c(2,4,8)) %>%
    mutate(temporal_threshold = paste0("temp_",temporal_threshold)) %>% 
    pivot_wider(id_cols=Group, names_from=temporal_threshold, values_from=transmission_prop) %>% 
    mutate(transmission_prop = sprintf("%.2f [%.2f-%.2f]", temp_4, temp_2, temp_8))
```



